---
title: "Dephosporylation"
author: "Nadine Prust"
date: "6/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dephosphorylation analysis
All necessary data is deposited on Pride and can be downloaded. The used files for this script can be found in the MQ_output_txt.zip file. Add the MQ_output_txt folder then to your project folder to access the required data. 
```{r}
library(ggplot2)
library(reshape2)
library(ggpubr)
```


Read list of synthetic peptides and extract p-site position 

```{r}

pArg.syn.peptides <- read.delim("MQ_output_txt/Synthetic_peptides/Dephosphorylation/Synthetic_peptides_lists/Syn_pArg_peptides.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
pArg.syn.peptides$positions <- lapply(gregexpr(pattern= "R\\(PO3\\)", pArg.syn.peptides$Sequence), min) 
pArg.syn.peptides$positions <- as.numeric(vapply(pArg.syn.peptides$positions, paste, collapse=", ", character(1L)))
pArg.syn.peptides$ID <- paste(pArg.syn.peptides$Peptide, pArg.syn.peptides$positions, sep = ",")

pST.syn.peptides <- read.delim("MQ_output_txt/Synthetic_peptides/Dephosphorylation/Synthetic_peptides_lists/Syn_pST_peptides.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
pST.syn.peptides$positions <-lapply(gregexpr(pattern= "[a-z]", pST.syn.peptides$Sequence), min) 
pST.syn.peptides$positions <- as.numeric(vapply(pST.syn.peptides$positions, paste, collapse = ", ", character(1L)))
pST.syn.peptides$ID <- paste(pST.syn.peptides$Pepetide, pST.syn.peptides$positions, sep =  ",")

pY.syn.peptides <- read.delim("MQ_output_txt/Synthetic_peptides/Dephosphorylation/Synthetic_peptides_lists/Syn_pY_peptides.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
pY.syn.peptides$positions <-lapply(gregexpr(pattern= "[a-z]", pY.syn.peptides$Sequence), min) 
pY.syn.peptides$positions <- as.numeric(vapply(pY.syn.peptides$positions, paste, collapse = ", ", character(1L)))
pY.syn.peptides$ID <- paste(pY.syn.peptides$Pepetide, pY.syn.peptides$positions, sep =  ",")


```

Read MaxQuant (v1.6.17.0) "Phospho (RHSTY)sites.txt" output for synthetic pSTY peptides

```{r}

load.input <- function(MQ.data){
  a <-read.delim(MQ.data, header = TRUE, stringsAsFactors = FALSE, dec = ".")
  a <- a[a$Potential.contaminant!= "+",]
  a <- a[which(a$Number.of.Phospho..RHSTY. != ""),]
  a$ID <- paste(a$Proteins, a$Positions.within.proteins, sep = "," )
  
  return(a)
}

#Ctrl samples

STY.Ctrl.1 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Ctrl/R1/txt/Phospho (RHSTY)Sites_STY_Ctrl1.txt")
STY.Ctrl.2 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Ctrl/R2/txt/Phospho (RHSTY)Sites_STY_Ctrl2.txt")
STY.Ctrl.3 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Ctrl/R3/txt/Phospho (RHSTY)Sites_STY_Ctrl3.txt")
STY.Ctrl.4 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Ctrl/R4/txt/Phospho (RHSTY)Sites_STY_Ctrl4.txt")

#extract correct pS/pT p-sites

ID.ST.Ctrl.1<- STY.Ctrl.1[STY.Ctrl.1$ID %in% pST.syn.peptides$ID ,]
ID.ST.Ctrl.1<- merge(ID.ST.Ctrl.1, pST.syn.peptides, by="ID")
ID.ST.Ctrl.2<- STY.Ctrl.2[STY.Ctrl.2$ID %in% pST.syn.peptides$ID ,]
ID.ST.Ctrl.2<- merge(ID.ST.Ctrl.2, pST.syn.peptides, by="ID")
ID.ST.Ctrl.3<- STY.Ctrl.3[STY.Ctrl.3$ID %in% pST.syn.peptides$ID ,]
ID.ST.Ctrl.3<- merge(ID.ST.Ctrl.3, pST.syn.peptides, by="ID")
ID.ST.Ctrl.4<- STY.Ctrl.4[STY.Ctrl.4$ID %in% pST.syn.peptides$ID ,]
ID.ST.Ctrl.4<- merge(ID.ST.Ctrl.4, pST.syn.peptides, by="ID")

ST.Ctrl <- cbind.data.frame("Condition" = "Ctrl","Rep" = "R1", "Number.id" = nrow(ID.ST.Ctrl.1), 
                            "PC.ST" = nrow(ID.ST.Ctrl.1)/nrow(pST.syn.peptides)*100)

ST.Ctrl <- rbind.data.frame(ST.Ctrl, 
                            cbind.data.frame("Condition" = "Ctrl", "Rep" = "R2", "Number.id" = nrow(ID.ST.Ctrl.2), 
                                             "PC.ST" = nrow(ID.ST.Ctrl.2)/nrow(pST.syn.peptides)*100))
ST.Ctrl <- rbind.data.frame(ST.Ctrl, 
                            cbind.data.frame("Condition" = "Ctrl", "Rep" = "R3", "Number.id" = nrow(ID.ST.Ctrl.3), 
                                             "PC.ST" = nrow(ID.ST.Ctrl.3)/nrow(pST.syn.peptides)*100))
ST.Ctrl <- rbind.data.frame(ST.Ctrl, 
                            cbind.data.frame("Condition" = "Ctrl", "Rep" = "R4", "Number.id" = nrow(ID.ST.Ctrl.4), 
                                             "PC.ST" = nrow(ID.ST.Ctrl.4)/nrow(pST.syn.peptides)*100))

#extract correct pY p-sites
ID.Y.Ctrl.1<- STY.Ctrl.1[STY.Ctrl.1$ID %in% pY.syn.peptides$ID ,]
ID.Y.Ctrl.1<- merge(ID.Y.Ctrl.1, pY.syn.peptides, by="ID")
ID.Y.Ctrl.2<- STY.Ctrl.2[STY.Ctrl.2$ID %in% pY.syn.peptides$ID ,]
ID.Y.Ctrl.2<- merge(ID.Y.Ctrl.2, pY.syn.peptides, by="ID")
ID.Y.Ctrl.3<- STY.Ctrl.3[STY.Ctrl.3$ID %in% pY.syn.peptides$ID ,]
ID.Y.Ctrl.3<- merge(ID.Y.Ctrl.3, pY.syn.peptides, by="ID")
ID.Y.Ctrl.4<- STY.Ctrl.4[STY.Ctrl.4$ID %in% pY.syn.peptides$ID ,]
ID.Y.Ctrl.4<- merge(ID.Y.Ctrl.4, pY.syn.peptides, by="ID")

Y.Ctrl <- cbind.data.frame("Condition" = "Ctrl","Rep" = "R1", "Number.id" = nrow(ID.Y.Ctrl.1), 
                           "PC.Y" = nrow(ID.Y.Ctrl.1)/nrow(pY.syn.peptides)*100)

Y.Ctrl <- rbind.data.frame(Y.Ctrl, 
                           cbind.data.frame("Condition" = "Ctrl", "Rep" = "R2", "Number.id" = nrow(ID.Y.Ctrl.2), 
                                            "PC.Y" = nrow(ID.Y.Ctrl.2)/nrow(pY.syn.peptides)*100))
Y.Ctrl <- rbind.data.frame(Y.Ctrl, 
                           cbind.data.frame("Condition" = "Ctrl", "Rep" = "R3", "Number.id" = nrow(ID.Y.Ctrl.3), 
                                            "PC.Y" = nrow(ID.Y.Ctrl.3)/nrow(pY.syn.peptides)*100))
Y.Ctrl <- rbind.data.frame(Y.Ctrl, 
                           cbind.data.frame("Condition" = "Ctrl", "Rep" = "R4", "Number.id" = nrow(ID.Y.Ctrl.4), 
                                            "PC.Y" = nrow(ID.Y.Ctrl.4)/nrow(pY.syn.peptides)*100))



#Stp1
STY.Stp1.1 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Stp1/R1/txt/Phospho (RHSTY)Sites_STY_Stp1_1.txt")
STY.Stp1.2 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Stp1/R2/txt/Phospho (RHSTY)Sites_STY_Stp1_2.txt")
STY.Stp1.3 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Stp1/R3/txt/Phospho (RHSTY)Sites_STY_Stp1_3.txt")
STY.Stp1.4 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/Stp1/R4/txt/Phospho (RHSTY)Sites_STY_Stp1_4.txt")


#extract correct pS/pT p-sites
ID.ST.Stp1.1<- STY.Stp1.1[STY.Stp1.1$ID %in% pST.syn.peptides$ID ,]
ID.ST.Stp1.1<- merge(ID.ST.Stp1.1, pST.syn.peptides, by="ID")
ID.ST.Stp1.2<- STY.Stp1.2[STY.Stp1.2$ID %in% pST.syn.peptides$ID ,]
ID.ST.Stp1.2<- merge(ID.ST.Stp1.2, pST.syn.peptides, by="ID")
ID.ST.Stp1.3<- STY.Stp1.3[STY.Stp1.3$ID %in% pST.syn.peptides$ID ,]
ID.ST.Stp1.3<- merge(ID.ST.Stp1.3, pST.syn.peptides, by="ID")
ID.ST.Stp1.4<- STY.Stp1.4[STY.Stp1.4$ID %in% pST.syn.peptides$ID ,]
ID.ST.Stp1.4<- merge(ID.ST.Stp1.4, pST.syn.peptides, by="ID")

ST.Stp1 <- cbind.data.frame("Condition" = "Stp1","Rep" = "R1", "Number.id" = nrow(ID.ST.Stp1.1), 
                            "PC.ST" = nrow(ID.ST.Stp1.1)/nrow(pST.syn.peptides)*100)

ST.Stp1 <- rbind.data.frame(ST.Stp1, 
                             cbind.data.frame("Condition" = "Stp1", "Rep" = "R2", "Number.id" = nrow(ID.ST.Stp1.2), 
                                              "PC.ST" = nrow(ID.ST.Stp1.2)/nrow(pST.syn.peptides)*100))
ST.Stp1 <- rbind.data.frame(ST.Stp1, 
                             cbind.data.frame("Condition" = "Stp1", "Rep" = "R3", "Number.id" = nrow(ID.ST.Stp1.3), 
                                              "PC.ST" = nrow(ID.ST.Stp1.3)/nrow(pST.syn.peptides)*100))
ST.Stp1 <- rbind.data.frame(ST.Stp1, 
                             cbind.data.frame("Condition" = "Stp1", "Rep" = "R4", "Number.id" = nrow(ID.ST.Stp1.4), 
                                              "PC.ST" = nrow(ID.ST.Stp1.4)/nrow(pST.syn.peptides)*100))
#Extract correct pY p-sites

ID.Y.Stp1.1<- STY.Stp1.1[STY.Stp1.1$ID %in% pY.syn.peptides$ID ,]
ID.Y.Stp1.1<- merge(ID.Y.Stp1.1, pY.syn.peptides, by="ID")
ID.Y.Stp1.2<- STY.Stp1.2[STY.Stp1.2$ID %in% pY.syn.peptides$ID ,]
ID.Y.Stp1.2<- merge(ID.Y.Stp1.2, pY.syn.peptides, by="ID")
ID.Y.Stp1.3<- STY.Stp1.3[STY.Stp1.3$ID %in% pY.syn.peptides$ID ,]
ID.Y.Stp1.3<- merge(ID.Y.Stp1.3, pY.syn.peptides, by="ID")
ID.Y.Stp1.4<- STY.Stp1.4[STY.Stp1.4$ID %in% pY.syn.peptides$ID ,]
ID.Y.Stp1.4<- merge(ID.Y.Stp1.4, pY.syn.peptides, by="ID")

Y.Stp1 <- cbind.data.frame("Condition" = "Stp1","Rep" = "R1", "Number.id" = nrow(ID.Y.Stp1.1), 
                            "PC.Y" = nrow(ID.Y.Stp1.1)/nrow(pY.syn.peptides)*100)

Y.Stp1 <- rbind.data.frame(Y.Stp1, 
                            cbind.data.frame("Condition" = "Stp1", "Rep" = "R2", "Number.id" = nrow(ID.Y.Stp1.2), 
                                             "PC.Y" = nrow(ID.Y.Stp1.2)/nrow(pY.syn.peptides)*100))
Y.Stp1 <- rbind.data.frame(Y.Stp1, 
                            cbind.data.frame("Condition" = "Stp1", "Rep" = "R3", "Number.id" = nrow(ID.Y.Stp1.3), 
                                             "PC.Y" = nrow(ID.Y.Stp1.3)/nrow(pY.syn.peptides)*100))
Y.Stp1 <- rbind.data.frame(Y.Stp1, 
                            cbind.data.frame("Condition" = "Stp1", "Rep" = "R4", "Number.id" = nrow(ID.Y.Stp1.4), 
                                             "PC.Y" = nrow(ID.Y.Stp1.4)/nrow(pY.syn.peptides)*100))


#rSAP

STY.rSAP.1 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/rSAP/R1/txt/Phospho (RHSTY)Sites_STY_rSAP1.txt")
STY.rSAP.2 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/rSAP/R2/txt/Phospho (RHSTY)Sites_STY_rSAP2.txt")
STY.rSAP.3 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/rSAP/R3/txt/Phospho (RHSTY)Sites_STY_rSAP3.txt")
STY.rSAP.4 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pSTY/rSAP/R4/txt/Phospho (RHSTY)Sites_STY_rSAP4.txt")


#Extract correct pS/pT p-sites
ID.ST.rSAP.1<- STY.rSAP.1[STY.rSAP.1$ID %in% pST.syn.peptides$ID ,]
ID.ST.rSAP.1<- merge(ID.ST.rSAP.1, pST.syn.peptides, by="ID")
ID.ST.rSAP.2<- STY.rSAP.2[STY.rSAP.2$ID %in% pST.syn.peptides$ID ,]
ID.ST.rSAP.2<- merge(ID.ST.rSAP.2, pST.syn.peptides, by="ID")
ID.ST.rSAP.3<- STY.rSAP.3[STY.rSAP.3$ID %in% pST.syn.peptides$ID ,]
ID.ST.rSAP.3<- merge(ID.ST.rSAP.3, pST.syn.peptides, by="ID")
ID.ST.rSAP.4<- STY.rSAP.4[STY.rSAP.4$ID %in% pST.syn.peptides$ID ,]
ID.ST.rSAP.4<- merge(ID.ST.rSAP.4, pST.syn.peptides, by="ID")

ST.rSAP <- cbind.data.frame("Condition" = "rSAP","Rep" = "R1", "Number.id" = nrow(ID.ST.rSAP.1), 
                            "PC.ST" = nrow(ID.ST.rSAP.1)/nrow(pST.syn.peptides)*100)

ST.rSAP <- rbind.data.frame(ST.rSAP, 
                            cbind.data.frame("Condition" = "rSAP", "Rep" = "R2", "Number.id" = nrow(ID.ST.rSAP.2), 
                                             "PC.ST" = nrow(ID.ST.rSAP.2)/nrow(pST.syn.peptides)*100))
ST.rSAP <- rbind.data.frame(ST.rSAP, 
                            cbind.data.frame("Condition" = "rSAP", "Rep" = "R3", "Number.id" = nrow(ID.ST.rSAP.3), 
                                             "PC.ST" = nrow(ID.ST.rSAP.3)/nrow(pST.syn.peptides)*100))
ST.rSAP <- rbind.data.frame(ST.rSAP, 
                            cbind.data.frame("Condition" = "rSAP", "Rep" = "R4", "Number.id" = nrow(ID.ST.rSAP.4), 
                                             "PC.ST" = nrow(ID.ST.rSAP.4)/nrow(pST.syn.peptides)*100))
#Extract correct pY p-sites

ID.Y.rSAP.1<- STY.rSAP.1[STY.rSAP.1$ID %in% pY.syn.peptides$ID ,]
ID.Y.rSAP.1<- merge(ID.Y.rSAP.1, pY.syn.peptides, by="ID")
ID.Y.rSAP.2<- STY.rSAP.2[STY.rSAP.2$ID %in% pY.syn.peptides$ID ,]
ID.Y.rSAP.2<- merge(ID.Y.rSAP.2, pY.syn.peptides, by="ID")
ID.Y.rSAP.3<- STY.rSAP.3[STY.rSAP.3$ID %in% pY.syn.peptides$ID ,]
ID.Y.rSAP.3<- merge(ID.Y.rSAP.3, pY.syn.peptides, by="ID")
ID.Y.rSAP.4<- STY.rSAP.4[STY.rSAP.4$ID %in% pY.syn.peptides$ID ,]
ID.Y.rSAP.4<- merge(ID.Y.rSAP.4, pY.syn.peptides, by="ID")

Y.rSAP <- cbind.data.frame("Condition" = "rSAP","Rep" = "R1", "Number.id" = nrow(ID.Y.rSAP.1), 
                           "PC.Y" = nrow(ID.Y.rSAP.1)/nrow(pY.syn.peptides)*100)

Y.rSAP <- rbind.data.frame(Y.rSAP, 
                           cbind.data.frame("Condition" = "rSAP", "Rep" = "R2", "Number.id" = nrow(ID.Y.rSAP.2), 
                                            "PC.Y" = nrow(ID.Y.rSAP.2)/nrow(pY.syn.peptides)*100))
Y.rSAP <- rbind.data.frame(Y.rSAP, 
                           cbind.data.frame("Condition" = "rSAP", "Rep" = "R3", "Number.id" = nrow(ID.Y.rSAP.3), 
                                            "PC.Y" = nrow(ID.Y.rSAP.3)/nrow(pY.syn.peptides)*100))
Y.rSAP <- rbind.data.frame(Y.rSAP, 
                           cbind.data.frame("Condition" = "rSAP", "Rep" = "R4", "Number.id" = nrow(ID.Y.rSAP.4), 
                                            "PC.Y" = nrow(ID.Y.rSAP.4)/nrow(pY.syn.peptides)*100))

```

Read MaxQuant "Phospho (RHSTY)sites.txt" output stored on the Pride archive for synthetic pArg peptides, add the respective prefix for the different replicates as stated below in the function

```{r}

#Ctrl

pArg.Ctrl.1 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Ctrl/R1/txt/Phospho (RHSTY)Sites_pArg_Ctrl1.txt")
pArg.Ctrl.2 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Ctrl/R2/txt/Phospho (RHSTY)Sites_pArg_Ctrl2.txt")
pArg.Ctrl.3 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Ctrl/R3/txt/Phospho (RHSTY)Sites_pArg_Ctrl3.txt")
pArg.Ctrl.4 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Ctrl/R4/txt/Phospho (RHSTY)Sites_pArg_Ctrl4.txt")

#Extract correct pArg peptides
ID.pArg.Ctrl.1 <- pArg.Ctrl.1[pArg.Ctrl.1$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.Ctrl.2 <- pArg.Ctrl.2[pArg.Ctrl.2$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.Ctrl.3 <- pArg.Ctrl.3[pArg.Ctrl.3$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.Ctrl.4 <- pArg.Ctrl.4[pArg.Ctrl.4$ID %in% pArg.syn.peptides$ID ,]

#Calculate percentage of identified synthetic peptides and combine replicates in one data frame 
pArg.ctrl <- cbind.data.frame("Condition" = "Ctrl","Rep" = "R1", "Number.id" = nrow(ID.pArg.Ctrl.1), "PC.pArg" = nrow(ID.pArg.Ctrl.1)/nrow(pArg.syn.peptides)*100)

pArg.ctrl <- rbind.data.frame(pArg.ctrl, 
                              cbind.data.frame("Condition" = "Ctrl", "Rep" = "R2", "Number.id" = nrow(ID.pArg.Ctrl.2), "PC.pArg" = nrow(ID.pArg.Ctrl.2)/nrow(pArg.syn.peptides)*100))
pArg.ctrl <- rbind.data.frame(pArg.ctrl, 
                              cbind.data.frame("Condition" = "Ctrl", "Rep" = "R3", "Number.id" = nrow(ID.pArg.Ctrl.3), "PC.pArg" = nrow(ID.pArg.Ctrl.3)/nrow(pArg.syn.peptides)*100))
pArg.ctrl <- rbind.data.frame(pArg.ctrl, 
                              cbind.data.frame("Condition" = "Ctrl", "Rep" = "R4", "Number.id" = nrow(ID.pArg.Ctrl.4), "PC.pArg" = nrow(ID.pArg.Ctrl.4)/nrow(pArg.syn.peptides)*100))

###Stp1


pArg.Stp1.1 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Stp1/R1/txt/Phospho (RHSTY)Sites_pArg_Stp1_1.txt")
pArg.Stp1.2 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Stp1/R2/txt/Phospho (RHSTY)Sites_pArg_Stp1_2.txt")
pArg.Stp1.3 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Stp1/R3/txt/Phospho (RHSTY)Sites_pArg_Stp1_3.txt")
pArg.Stp1.4 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/Stp1/R4/txt/Phospho (RHSTY)Sites_pArg_Stp1_4.txt")

#Extract correct pArg peptides
ID.pArg.Stp1.1 <- pArg.Stp1.1[pArg.Stp1.1$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.Stp1.2 <- pArg.Stp1.2[pArg.Stp1.2$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.Stp1.3 <- pArg.Stp1.3[pArg.Stp1.3$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.Stp1.4 <- pArg.Stp1.4[pArg.Stp1.4$ID %in% pArg.syn.peptides$ID ,]

#Calculate percentage of identified synthetic peptides and combine replicates in one data frame 
pArg.Stp1 <- cbind.data.frame("Condition" = "Stp1","Rep" = "R1", "Number.id" = nrow(ID.pArg.Stp1.1), "PC.pArg" = nrow(ID.pArg.Stp1.1)/nrow(pArg.syn.peptides)*100)

pArg.Stp1 <- rbind.data.frame(pArg.Stp1, 
                              cbind.data.frame("Condition" = "Stp1","Rep" = "R2", "Number.id" = nrow(ID.pArg.Stp1.2), "PC.pArg" = nrow(ID.pArg.Stp1.2)/nrow(pArg.syn.peptides)*100))
pArg.Stp1 <- rbind.data.frame(pArg.Stp1, 
                              cbind.data.frame("Condition" = "Stp1","Rep" = "R3", "Number.id" = nrow(ID.pArg.Stp1.3), "PC.pArg" = nrow(ID.pArg.Stp1.3)/nrow(pArg.syn.peptides)*100))
pArg.Stp1 <- rbind.data.frame(pArg.Stp1, 
                              cbind.data.frame("Condition" = "Stp1","Rep" = "R4", "Number.id" = nrow(ID.pArg.Stp1.4), "PC.pArg" = nrow(ID.pArg.Stp1.4)/nrow(pArg.syn.peptides)*100))

####rSAP


pArg.rSAP1 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/rSAP/R1/txt/Phospho (RHSTY)Sites_pArg_rSAP1.txt")
pArg.rSAP2 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/rSAP/R2/txt/Phospho (RHSTY)Sites_rSAP_2.txt")
pArg.rSAP3 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/rSAP/R3/txt/Phospho (RHSTY)Sites_rSAP_3.txt")
pArg.rSAP4 <- load.input("MQ_output_txt/Synthetic_peptides/Dephosphorylation/pArg/rSAP/R4/txt/Phospho (RHSTY)Sites_rSAP_4.txt")

#Extract correct pArg peptides
ID.pArg.rSAP.1 <- pArg.rSAP1[pArg.rSAP1$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.rSAP.2 <- pArg.rSAP2[pArg.rSAP2$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.rSAP.3 <- pArg.rSAP3[pArg.rSAP3$ID %in% pArg.syn.peptides$ID ,]
ID.pArg.rSAP.4 <- pArg.rSAP4[pArg.rSAP4$ID %in% pArg.syn.peptides$ID ,]

#Calculate percentage of identified synthetic peptides and combine replicates in one data frame 
pArg.rSAP <- cbind.data.frame("Condition" = "rSAP", "Rep" = "R1", "Number.id" = nrow(ID.pArg.rSAP.1), "PC.pArg" = nrow(ID.pArg.rSAP.1)/nrow(pArg.syn.peptides)*100)

pArg.rSAP <- rbind.data.frame(pArg.rSAP, 
                              cbind.data.frame("Condition" = "rSAP","Rep" = "R2", "Number.id" = nrow(ID.pArg.rSAP.2), "PC.pArg" = nrow(ID.pArg.rSAP.2)/nrow(pArg.syn.peptides)*100))
pArg.rSAP <- rbind.data.frame(pArg.rSAP, 
                              cbind.data.frame("Condition" = "rSAP","Rep" = "R3", "Number.id" = nrow(ID.pArg.rSAP.3), "PC.pArg" = nrow(ID.pArg.rSAP.3)/nrow(pArg.syn.peptides)*100))
pArg.rSAP <- rbind.data.frame(pArg.rSAP, 
                              cbind.data.frame("Condition" = "rSAP","Rep" = "R4", "Number.id" = nrow(ID.pArg.rSAP.4), "PC.pArg" = nrow(ID.pArg.rSAP.4)/nrow(pArg.syn.peptides)*100))

```

Plot Boxplots with percentages of identified synthetic peptides

```{r}
pST <- rbind.data.frame(ST.Ctrl, ST.Stp1, ST.rSAP)
pY <- rbind.data.frame(Y.Ctrl, Y.Stp1, Y.rSAP)
pArg <- rbind.data.frame(pArg.ctrl, pArg.Stp1, pArg.rSAP)

pST.plot <- 
  pST %>%
  ggplot(aes(x=Condition, y=PC.ST, fill=Condition))+
  geom_boxplot(show.legend = FALSE )+
  scale_fill_manual(values = c("Ctrl" = "#80146E", "Stp1" = "#6771B8", "rSAP" = "#58C9BC") )+
  stat_summary(fun=mean, geom="point", shape=23, size=3, show.legend = FALSE)+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line())+
  ylab("Identified synthetic pST peptides [%]")

pY.plot <- 
  pY %>%
  ggplot(aes(x=Condition, y=PC.Y, fill=Condition))+
  geom_boxplot(show.legend = FALSE  )+
  scale_fill_manual(values = c("Ctrl" = "#80146E", "Stp1" = "#6771B8", "rSAP" = "#58C9BC") )+
  stat_summary(fun=mean, geom="point", shape=23, size=3, show.legend = FALSE)+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line())+
  ylab("Identified synthetic pY peptides [%]")

pArg.plot <- pArg %>%
ggplot(aes(x=Condition, y=PC.pArg, fill=Condition))+
  geom_boxplot(show.legend = FALSE)+
  scale_fill_manual(values = c("Ctrl" = "#80146E", "Stp1" = "#6771B8", "rSAP" = "#58C9BC") )+
  stat_summary(fun=mean, geom="point", shape=23, size=3, show.legend = FALSE)+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line())+
  ylab("Identified synthetic pArg peptides [%]")

```

## Dephosphorylation of endogenous peptides

Read MaxQuant "Evidence.txt" output from endogenous dephosphorylated samples

```{r}
evidence.Ctrl.1 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Ctrl/R1/txt/evidence_Ctrl1.txt" , header = TRUE, stringsAsFactors = FALSE, dec = ".")
evidence.Ctrl.2 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Ctrl/R2/txt/evidence_Ctrl2.txt" , header = TRUE, stringsAsFactors = FALSE, dec = ".")
evidence.Ctrl.3 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Ctrl/R3/txt/evidence_Ctrl3.txt" , header = TRUE, stringsAsFactors = FALSE, dec = ".")

evidence.Stp1.1<- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Stp1/R1/txt/evidence_Stp1_1.txt" , header = TRUE, stringsAsFactors = FALSE, dec = ".")
evidence.Stp1.2<- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Stp1/R2/txt/evidence_Stp1_2.txt" , header = TRUE, stringsAsFactors = FALSE, dec = ".")
evidence.Stp1.3<- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Stp1/R3/txt/evidence_Stp1_3.txt" , header = TRUE, stringsAsFactors = FALSE, dec = ".")
```

Filter data for "Reverse" and "Potential Contaminants" as well ans on duplicated entries on "Modified Sequence" and calculate the percentage of phosphopeptides

```{r}
average_numbers <- function(df1, df2, df3){
  a <- df1[df1$Reverse != "+",]
  a<- df1[df1$Potential.contaminant != "+",]
  a<- df1[!duplicated(df1$Modified.sequence),]
  
  b <- df2[df2$Reverse != "+",]
  b<- df2[df2$Potential.contaminant != "+",]
  b<- df2[!duplicated(df2$Modified.sequence),]
  
  c <- df3[df3$Reverse != "+",]
  c<- df3[df3$Potential.contaminant != "+",]
  c<- df3[!duplicated(df3$Modified.sequence),]
  
  d<- rbind.data.frame(cbind( "No.peptides" = nrow(a), 
                        "No.phosphopeptides" = nrow(a[a$Phospho..RHSTY. != 0,]), 
                        "PC.ppeptides" = nrow(a[a$Phospho..RHSTY. != 0,]) / nrow(a)*100),
                       cbind( "No.peptides" = nrow(b), 
                                         "No.phosphopeptides" = nrow(b[b$Phospho..RHSTY. != 0,]), 
                                         "PC.ppeptides" =nrow(b[b$Phospho..RHSTY. != 0,]) / nrow(b)*100),
                       cbind("No.peptides" = nrow(c), 
                                         "No.phosphopeptides" = nrow(c[c$Phospho..RHSTY. != 0,]), 
                                         "PC.ppeptides" =nrow(c[c$Phospho..RHSTY. != 0,]) / nrow(c)*100))

  return(d)
}

Ctrl.peptides <- average_numbers(evidence.Ctrl.1, evidence.Ctrl.2, evidence.Ctrl.3)
STp1.peptides <- average_numbers(evidence.Stp1.1, evidence.Stp1.2, evidence.Stp1.3)
```

Calculate the average of identified peptides, phosphopeptides in percentage of identified phosphopeptides

```{r}
Peptide.number.comp <- rbind.data.frame(cbind("Sample"= "Ctrl","ID" = "Peptides", "Average" = mean(Ctrl.peptides$No.peptides), 
                                              "SD" =sd(Ctrl.peptides$No.peptides)),
                                        cbind("Sample" = "Ctrl", "ID" = "Phosphopeptides", 
                                              "Average" = mean(Ctrl.peptides$No.phosphopeptides), 
                                              "SD" =sd(Ctrl.peptides$No.phosphopeptides)),
                                        cbind("Sample"= "Stp1","ID" = "Peptides", "Average" = mean(STp1.peptides$No.peptides), 
                                              "SD" = sd(STp1.peptides$No.peptides)),
                                        cbind("Sample" = "Stp1", "ID" = "Phosphopeptides", 
                                              "Average" = mean(STp1.peptides$No.phosphopeptides), 
                                              "SD" = sd(STp1.peptides$No.phosphopeptides)))
Peptide.number.comp$Average <- as.numeric(as.character(Peptide.number.comp$Average))
Peptide.number.comp$SD <- as.numeric(as.character(Peptide.number.comp$SD))
```

Plot bar graph with number average number of identification

```{r}
Peptide.plot <- 
  ggplot(Peptide.number.comp, aes(x=Sample, y= Average, fill= Sample))+
  geom_bar( stat="identity", position = position_dodge())+
  geom_errorbar(data = Peptide.number.comp, aes(x=Sample, ymin=Average-SD, ymax=Average+SD), width=.1) +
  scale_fill_manual(values = c("Ctrl" = "#80146E", "Stp1" = "#6771B8"))+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line())+
  ylab("Average number of identification")+
    facet_wrap(~ID)
```

## Phosphosite distribution analysis

Read MaxQuant "Phospho (RHSTY)Sites.txt" output

```{r}
  Ctrl.1 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Ctrl/R1/txt/Phospho (RHSTY)Sites_Ctrl1.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
  Ctrl.2 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Ctrl/R2/txt/Phospho (RHSTY)Sites_Ctrl2.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
  Ctrl.3 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Ctrl/R3/txt/Phospho (RHSTY)Sites_Ctrl3.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
  Stp1.1 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Stp1/R1/txt/Phospho (RHSTY)Sites_Stp1_1.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
  Stp1.2 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Stp1/R2/txt/Phospho (RHSTY)Sites_Stp1_2.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
  Stp1.3 <- read.delim("MQ_output_txt/Endogenous_data/Dephosphorylation/Stp1/R3/txt/Phospho (RHSTY)Sites_Stp1_3.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
```  

Filter for "Reverse", and "Potential contaminants" and determine the average number and fractions of pS. pT, pY, pH and pR 

```{r}
psites.extraction <- function(df1, df2, df3, SAMPLE){
    a <- df1[df1$Reverse != "+",]
    a<- a[a$Potential.contaminant != "+",]
    
    a.dis <- data.frame(
      category=as.character(c("pS", "pT", "pY", "pH", "pR")),
      count=as.numeric(as.character(c(nrow(a[a$Amino.acid == "S",]), nrow(a[a$Amino.acid == "T",]), 
              nrow(a[a$Amino.acid == "Y",]), nrow(a[a$Amino.acid == "H",]),
              nrow(a[a$Amino.acid == "R",])))),
      Replicate = "R1"
    )
    a.dis$fraction <- a.dis$count/sum(a.dis$count)*100
    
    b <- df2[df2$Reverse != "+",]
    b<- b[b$Potential.contaminant != "+",]
    
    b.dis <- data.frame(
      category=as.character(c("pS", "pT", "pY", "pH", "pR")),
      count= as.numeric(as.character(c(nrow(b[b$Amino.acid == "S",]), nrow(b[b$Amino.acid == "T",]), 
              nrow(b[b$Amino.acid == "Y",]), nrow(b[b$Amino.acid == "H",]),
              nrow(b[b$Amino.acid == "R",])))),
      Replicate = "R2"
    )
    b.dis$fraction <- b.dis$count/ sum(b.dis$count)*100
    
    c <- df3[df3$Reverse != "+",]
    c<- c[c$Potential.contaminant != "+",]
    
    c.dis <- data.frame(
      category=as.character(c("pS", "pT", "pY", "pH", "pR")),
      count=as.numeric(as.character(c(nrow(c[c$Amino.acid == "S",]), nrow(c[c$Amino.acid == "T",]), 
              nrow(c[c$Amino.acid == "Y",]), nrow(c[c$Amino.acid == "H",]),
              nrow(c[c$Amino.acid == "R",])))),
      Replicate = "R3"
    )
    c.dis$fraction <- c.dis$count / sum(c.dis$count)*100
  d <- rbind.data.frame(a.dis, b.dis, c.dis)
      d<- d %>% group_by(category) %>% mutate("Average" = mean(count), "SD" = sd(count), 
                                              "Avg.fraction" = mean(fraction), "SD.fraction" = sd(fraction))
    
    d<- d[!duplicated(d$category), c("category", "Average", "SD", "Avg.fraction", "SD.fraction")]
    d$Sample <- SAMPLE
    return(d)
    
  }
  
  Ctrl.psites <- psites.extraction(Ctrl.1, Ctrl.2, Ctrl.3, "Ctrl")
  #Adjust values for stack error bars
  Ctrl.psites$vadj <- ifelse(Ctrl.psites$category == "pY",
                             0,
                             ifelse(Ctrl.psites$category == "pT",
                                    Ctrl.psites[Ctrl.psites$category == "pY", "Avg.fraction"],
                                    ifelse(Ctrl.psites$category =="pS",
                                           sum(Ctrl.psites[Ctrl.psites$category == "pY", "Avg.fraction"], 
                                               Ctrl.psites[Ctrl.psites$category == "pT", "Avg.fraction"]),
                                           ifelse(Ctrl.psites$category == "pR", 
                                                  sum(Ctrl.psites[Ctrl.psites$category == "pY", "Avg.fraction"], 
                                                      Ctrl.psites[Ctrl.psites$category == "pT", "Avg.fraction"],
                                                      Ctrl.psites[Ctrl.psites$category == "pS", "Avg.fraction"]),
                                                  sum(Ctrl.psites[Ctrl.psites$category == "pY", "Avg.fraction"], 
                                                      Ctrl.psites[Ctrl.psites$category == "pT", "Avg.fraction"],
                                                      Ctrl.psites[Ctrl.psites$category == "pS", "Avg.fraction"],
                                                      Ctrl.psites[Ctrl.psites$category == "pR", "Avg.fraction"]))))
)
  
  Ctrl.psites$vadj <-  vapply(Ctrl.psites$vadj, paste, collapse = ", ", character(1L))
 Stp1.psites <- psites.extraction(Stp1.1, Stp1.2, Stp1.3, "Stp1")
 
 Stp1.psites$vadj <- ifelse(Stp1.psites$category == "pY",
                            0,
                            ifelse(Stp1.psites$category == "pT",
                                   Stp1.psites[Stp1.psites$category == "pY", "Avg.fraction"],
                                   ifelse(Stp1.psites$category =="pS",
                                          sum(Stp1.psites[Stp1.psites$category == "pY", "Avg.fraction"], 
                                              Stp1.psites[Stp1.psites$category == "pT", "Avg.fraction"]),
                                          ifelse(Stp1.psites$category == "pR", 
                                                 sum(Stp1.psites[Stp1.psites$category == "pY", "Avg.fraction"], 
                                                     Stp1.psites[Stp1.psites$category == "pT", "Avg.fraction"],
                                                     Stp1.psites[Stp1.psites$category == "pS", "Avg.fraction"]),
                                                 sum(Stp1.psites[Stp1.psites$category == "pY", "Avg.fraction"], 
                                                     Stp1.psites[Stp1.psites$category == "pT", "Avg.fraction"],
                                                     Stp1.psites[Stp1.psites$category == "pS", "Avg.fraction"],
                                                     Stp1.psites[Stp1.psites$category == "pR", "Avg.fraction"]))))
 )
 Stp1.psites$vadj <-  vapply(Stp1.psites$vadj, paste, collapse = ", ", character(1L))
 
 psites <- rbind.data.frame(Ctrl.psites, Stp1.psites)
 psites$vadj <- as.numeric(psites$vadj)
```

Plot stacked bar graph with percentages of p-sites

```{r}
psite.dis <- 
   ggplot(psites, aes(x=Sample, y=Avg.fraction, fill=category))+
   geom_bar(stat = "identity", position = "stack")+
   geom_errorbar(data = psites, aes(x=Sample, ymin= Avg.fraction+vadj - SD.fraction, ymax=Avg.fraction+vadj+SD.fraction),
                 stat = "identity", position = "identity", width=.1)+
     geom_text(data = psites, aes(x=Sample, y= Avg.fraction, label=Average), 
               position = position_stack(vjust = 0.5))+
   scale_fill_manual(values = c("pS"= "#b3cde3","pT"= "#8c96c6","pY"= "#8856a7" ,"pH"= "#810f7c","pR"= "#210320"))+
   theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line())+
   ylab("Avergage number of p-sites [%]")
```

Combine all graphs in one figure

```{r}

ggarrange(ggarrange(pArg.plot, pST.plot, pY.plot, labels = c("a)", "b)", "c)"), ncol = 3),
          ggarrange(Peptide.plot, psite.dis, labels = c("d)", "e)"), ncol=2),
          nrow = 2)
```
