---
title: Andromeda localization probability and Andromeda score comaprison for pArg and pHSTY sites
  peptides
author: "Nadine Prust"
date: "6/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comparison of Andromeda localization probability and Andromeda score for endogenous peptides that were either exclusively identified being phosphorylated on arginine that were also identified  to be phosphorylated on H,S,T, or Y. and pArg peptides 

All necessary data is deposited on Pride and can be downloaded. The used files for this script can be found in the MQ_output_txt.zip file. Add the MQ_output_txt folder then to your project folder to access the required data.

```{r}
 library(stringr)
library(reshape2)
```

Read Phospho (RHSTY).txt from combined MaxQuant (v1.6.3.4) search and read uniprot list containing protein names

```{r}

phospho_RHSTY <- read.delim("MQ_output_txt/Endogenous_data/LFQ_Saureus/txt/Phospho (RHSTY)Sites_exp7.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
uniprotname <- read.delim("MQ_output_txt/Endogenous_data/LFQ_Saureus/txt/uniprot_list_pRHSTY.txt", header = TRUE, stringsAsFactors = FALSE)
```

Extract important columns and filter data

```{r}
pRHSTY.aureus<- phospho_RHSTY[, c("Protein","Amino.acid", "Positions.within.proteins", "Number.of.Phospho..RHSTY.", "Phospho..RHSTY..Probabilities",
                                        "Intensity.SdrE_1", "Intensity.SdrE_2", "Intensity.SdrE_3", "Intensity.SdrE_4",
                                        "Intensity.Stk1_1", "Intensity.Stk1_2", "Intensity.Stk1_3", "Intensity.Stk1_4",
                                        "Intensity.Stp1_1", "Intensity.Stp1_2", "Intensity.Stp1_3", "Intensity.Stp1_4", "Delta.score", "Score",
                                        "Localization.prob", "Sequence.window",
                                        "Reverse", "Potential.contaminant",
                                        "Localization.prob.SdrE_1","Localization.prob.SdrE_2", "Localization.prob.SdrE_3", "Localization.prob.SdrE_4", 
                                        "Localization.prob.Stk1_1", "Localization.prob.Stk1_2", "Localization.prob.Stk1_3", "Localization.prob.Stk1_4",
                                        "Localization.prob.Stp1_1", "Localization.prob.Stp1_2", "Localization.prob.Stp1_3", "Localization.prob.Stp1_4",
                                        "Best.localization.evidence.ID", "Best.localization.raw.file", "Best.localization.scan.number"
                                        
)]
# removing reversed and contaminants


pRHSTY.aureus<- pRHSTY.aureus[pRHSTY.aureus$Reverse != "+",]
pRHSTY.aureus <- pRHSTY.aureus[pRHSTY.aureus$Potential.contaminant != "+",]
pRHSTY.aureus$ID <- paste(pRHSTY.aureus$Protein, pRHSTY.aureus$Amino.acid, pRHSTY.aureus$Positions.within.proteins,
                        sep = ",")

pRHSTY.aureus$Sequence <- gsub("\\(\\d*\\.?\\d*\\)","", pRHSTY.exp7$Phospho..RHSTY..Probabilities)
pRHSTY.aureus <- pRHSTY.aureus[pRHSTY.aureus$Number.of.Phospho..RHSTY. != "",]

```

Perform log2 transformation and median normalization of the intensities

```{r}
pRHSTY.aureus[,grep("Intensity.*", colnames(pRHSTY.aureus))] <- apply(pRHSTY.aureus[,grep("Intensity.*", colnames(pRHSTY.aureus))], 2, FUN=log2 )
# change -Inf to NA
# take all columns containing LFQ in their header
intensitycolumns <- grep("Intensity.*", colnames(pRHSTY.aureus))
# apply log transformation on all columns containing LFQ in their column name
pRHSTY.aureus[,intensitycolumns] <- apply(pRHSTY.aureus[, intensitycolumns], 2, function(x) { x[is.infinite(x)] <- NA; x})  

### median normalization - calculate median per sample (on column) and substract it from each row

pRHSTY.aureus[,intensitycolumns] <- apply(pRHSTY.aureus[, intensitycolumns], 2, function(x) {x- median(x, na.rm = TRUE)}) 

### calculate avg

pRHSTY.aureus$SdrE <- rowMeans(pRHSTY.aureus[,grep("Intensity.SdrE", colnames(pRHSTY.aureus))], na.rm = TRUE)
pRHSTY.aureus$Stk1 <- rowMeans(pRHSTY.aureus[,grep("Intensity.Stk1", colnames(pRHSTY.aureus))], na.rm = TRUE)
pRHSTY.aureus$Stp1 <- rowMeans(pRHSTY.aureus[,grep("Intensity.Stp1", colnames(pRHSTY.aureus))], na.rm = TRUE)

###count NA per replicate

pRHSTY.aureus$NA_SdrE <- apply(pRHSTY.aureus[,grep("Intensity.SdrE", colnames(pRHSTY.aureus))],1,function(x) sum(is.na(x)))

pRHSTY.aureus$NA_Stk1 <- apply(pRHSTY.aureus[,grep("Intensity.Stk1", colnames(pRHSTY.aureus))],1,function(x) sum(is.na(x)))

pRHSTY.aureus$NA_Stp1 <- apply(pRHSTY.aureus[,grep("Intensity.Stp1", colnames(pRHSTY.aureus))],1,function(x) sum(is.na(x)))

```

Filter data for valid values: 3x in at least one group 

```{r}
pRHSTY.aureus.filtered <- pRHSTY.aureus[(pRHSTY.aureus$NA_SdrE <=1 | pRHSTY.aureus$NA_Stk1 <=1 | pRHSTY.aureus$NA_Stp1),]

```

Determine sequence without missed cleavages to take higher amount of missed cleaveged due to pArg into account. 

```{r}
pArg.aureus <- pRHSTY.aureus.filtered[pRHSTY.aureus.filtered$Amino.acid == "R",]
pArg.aureus$missed.cleavages <- str_count(pArg.aureus$Sequence, "K|R")-1
pArg.aureus$before.missed.cleavage <- str_extract(pArg.aureus$Sequence, "^(.*?R|K)")
pArg.aureus$after.missed.cleavage <- str_extract(pArg.aureus$Sequence, "R.*|K.*")
pArg.aureus.short <- pArg.aureus %>% select(Protein, Amino.acid, Positions.within.proteins, ID, Number.of.Phospho..RHSTY., Localization.prob, 
                                            Best.localization.evidence.ID, Best.localization.raw.file, Best.localization.scan.number, Sequence.window,
                                            Sequence, before.missed.cleavage, after.missed.cleavage)



pHSTY.aureus <- pRHSTY.aureus.filtered[pRHSTY.aureus.filtered$Amino.acid != "R",]
pHSTY.aureus$missed.cleavages <- str_count(pHSTY.aureus$Sequence, "K|R")-1
pHSTY.aureus$before.missed.cleavage <- pHSTY.aureus$Sequence
pHSTY.aureus$after.missed.cleavage <- pHSTY.aureus$Sequence
pHSTY.aureus.short <- pHSTY.aureus %>% select(Protein, Amino.acid, Positions.within.proteins,ID, Number.of.Phospho..RHSTY., Localization.prob, 
                                              Best.localization.evidence.ID, Best.localization.raw.file, Best.localization.scan.number, Sequence.window,
                                              Sequence, before.missed.cleavage, after.missed.cleavage)

sequence.comparison <- inner_join(pArg.aureus.short, pHSTY.aureus.short, by="Sequence")
sequence.comparison <- sequence.comparison[,-23]

sequence.comparison.before.missed.cleavage <- inner_join(pArg.aureus.short, pHSTY.aureus.short, by="before.missed.cleavage")
sequence.comparison.before.missed.cleavage <- sequence.comparison.before.missed.cleavage[,-23]
sequence.comparison.after.missed.cleavage <- inner_join(pArg.aureus.short, pHSTY.aureus.short, by="after.missed.cleavage")
sequence.comparison.after.missed.cleavage <- sequence.comparison.after.missed.cleavage[,-23]
names.comparison <- colnames(sequence.comparison.before.missed.cleavage)
colnames(sequence.comparison) <- names.comparison
colnames(sequence.comparison.after.missed.cleavage) <- names.comparison

combined <- rbind.data.frame(sequence.comparison, sequence.comparison.before.missed.cleavage, sequence.comparison.after.missed.cleavage)
combined.unique <- combined[which(!duplicated(combined$ID.x)),]
combined.doubly.phosphorylated <- combined[combined$Number.of.Phospho..RHSTY..x >= 2,]

identifier <- c(combined$ID.x, combined$ID.y)
test2 <-as.data.frame(gsub(",.*", "", identifier$`combined$ID.x`))

multiple.ppeptide <- pRHSTY.aureus[pRHSTY.aureus$ID %in% identifier,]
multiple.ppeptide$ID.aa <- ifelse(multiple.ppeptide$Amino.acid == "R", "R", "HSTY")

pRHSTY.aureus.filtered$multiple <- ifelse(pRHSTY.aureus.filtered$ID %in% identifier, "Yes", "NO")
pRHSTY.aureus.filtered$ID.aa <- ifelse(pRHSTY.aureus.filtered$Amino.acid == "R", "R", "HSTY")

#filter for localization prob of >= 0.75

classI.combined <- combined[combined$Localization.prob.x >= 0.75,]
classI.identifier <- c(classI.combined$ID.x, classI.combined$ID.y)
classI.multiple.ppeptides <- pRHSTY.aureus[pRHSTY.aureus$ID %in% classI.identifier,]

both.classI.combined <- combined[combined$Localization.prob.x >= 0.75 & combined$Localization.prob.y >= 0.75,]
both.classI.combined <- both.classI.combined[which(!duplicated(both.classI.combined$ID.x)),]
```

Plot Andromeda Score for pArg and pHSTY p-sites

```{r}

Score.plot <-
pRHSTY.aureus.filtered %>%
  mutate(ID.aa = factor(ID.aa, levels = c("HSTY" ,"R"))) %>%
  mutate(multiple = factor(multiple, levels = c("Yes" ,"NO"))) %>%
ggplot(aes(x=ID.aa, y=Score, fill=ID.aa)) +
  geom_violin(color="black", show.legend = FALSE)+
  scale_fill_manual(values = c("#01665e","#762a83")) +
  geom_boxplot(width=0.1, show.legend = FALSE)+
  stat_summary(fun=mean, geom="point", shape=23, size=3, show.legend = FALSE)+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line())+
  ylab("Andromeda score")+
  facet_wrap(~multiple)
```

Determine the Andromeda localization probability distribution of pArg and pHSTY peptides

```{r}
breaks.num <- c(0.4,0.5,0.6,0.7,0.75,0.8,0.85,0.9,0.95,0.99,1)


localization.analysis <- function(df, loc, breaks.num,amino.acid1, amino.acid2, multiple1, multiple2){
  a <- df[df$ID.aa == amino.acid1 & df$multiple == multiple1,]
  a <- 
    a %>% pull(loc) %>% hist(breaks = breaks.num)
  data.a <- as.data.frame(breaks.num)
  data.a <- cbind.data.frame(data.a[-1,], a$counts)
  
  total.number <- sum(data.a[,2])
  data.a$pc <- data.a[,2]/total.number*100
  colnames(data.a) <-c("MQ.localization.probability", "count", "pc.peptides")
  
  
  data.a$peptide.pc <- data.a$pc.peptides
  penu <- nrow(data.a)
  for(i in 2:penu){
    data.a$peptide.pc[i] <- {data.a$peptide.pc[i] + data.a$peptide.pc[i-1]}
    
  }
  data.a$AA <- amino.acid1
  data.a$Multiple.ID <- multiple1
  
  b <- df[df$ID.aa == amino.acid1 & df$multiple == multiple2,]
  b <- 
    b %>% pull(loc) %>% hist(breaks = breaks.num)
  data.b <- as.data.frame(breaks.num)
  data.b <- cbind.data.frame(data.b[-1,], b$counts)
  
  total.number <- sum(data.b[,2])
  data.b$pc <- data.b[,2]/total.number*100
  colnames(data.b) <-c("MQ.localization.probability", "count", "pc.peptides")
  
  
  data.b$peptide.pc <- data.b$pc.peptides
  penu <- nrow(data.b)
  for(i in 2:penu){
    data.b$peptide.pc[i] <- {data.b$peptide.pc[i] + data.b$peptide.pc[i-1]}
    
  }
  data.b$AA <- amino.acid1
  data.b$Multiple.ID <- multiple2
  
  
  c <- df[df$ID.aa == amino.acid2 & df$multiple == multiple1,]
  c <- 
    c %>% pull(loc) %>% hist(breaks = breaks.num)
  data.c <- as.data.frame(breaks.num)
  data.c <- cbind.data.frame(data.c[-1,], c$counts)
  
  total.number <- sum(data.c[,2])
  data.c$pc <- data.c[,2]/total.number*100
  colnames(data.c) <-c("MQ.localization.probability", "count", "pc.peptides")
  
  
  data.c$peptide.pc <- data.c$pc.peptides
  penu <- nrow(data.c)
  for(i in 2:penu){
    data.c$peptide.pc[i] <- {data.c$peptide.pc[i] + data.c$peptide.pc[i-1]}
    
  }
  data.c$AA <- amino.acid2
  data.c$Multiple.ID <- multiple1
  
  
  d <- df[df$ID.aa == amino.acid2 & df$multiple == multiple2,]
  d <- 
    d %>% pull(loc) %>% hist(breaks = breaks.num)
  data.d <- as.data.frame(breaks.num)
  data.d <- cbind.data.frame(data.d[-1,], d$counts)
  
  total.number <- sum(data.d[,2])
  data.d$pc <- data.d[,2]/total.number*100
  colnames(data.d) <-c("MQ.localization.probability", "count", "pc.peptides")
  
  
  data.d$peptide.pc <- data.d$pc.peptides
  penu <- nrow(data.d)
  for(i in 2:penu){
    data.d$peptide.pc[i] <- {data.d$peptide.pc[i] + data.d$peptide.pc[i-1]}
    
  }
  data.d$AA <- amino.acid2
  data.d$Multiple.ID <- multiple2
  
  data.combined <- rbind(data.a, data.b, data.c, data.d)
  return(data.combined)
  
}



Localization <- localization.analysis(pRHSTY.aureus.filtered,"Localization.prob", breaks.num, "HSTY", "R", "Yes", "NO" )

```

Plot Andromeda localization probaility distribution

```{r}

localization.plot <-
Localization %>% arrange(AA) %>%
  mutate(AA = factor(AA, levels = c("HSTY","R"))) %>%
  mutate(Multiple.ID = factor(Multiple.ID, levels = c("Yes","NO"))) %>%
  ggplot(aes(x=peptide.pc, y=MQ.localization.probability, color= AA))+
  geom_line(aes(linetype=Multiple.ID), size=1)+
  scale_color_manual(values=c("#01665e","#762a83"))+
  xlab("% of identified peptides")+
  ylab("MaxQuant localization probability")+
  theme(panel.background = element_blank(), axis.line = element_line(), legend.position = c(0.8, 0.5))

```

Combin plots

```{r}
ggarrange(localization.plot, Score.plot, labels = c("a)", "b)"), ncol = 2, nrow = 1)
```

