---
title: "Missed cleavages"
author: "Nadine Prust"
date: "10/28/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analysis of missed cleavages of pArg, pHis and pSTY 

All necessary data is deposited on Pride and can be downloaded. The used files for this script can be found in the MQ_output_txt.zip file. Add the MQ_output_txt folder then to your project folder to access the required data. 

```{r}
library(stringr)
library(dplyr)
library(ggplot2)
library(forcats)
library(colorspace)

```

Read "Phospho (RHSTY).txt" from combined search of all three mutants with MaxQuant (v1.6.3.4)

```{r}
pRHSTY.exp7 <- read.delim("MQ_output_txt/Endogenous_data/LFQ_Saureus/txt/Phospho (RHSTY)Sites_exp7.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
```

Filter for contaminants and reversed

```{r}
pRHSTY.exp7 <- pRHSTY.exp7[pRHSTY.exp7$Potential.contaminant != "+",]
pRHSTY.exp7 <- pRHSTY.exp7[pRHSTY.exp7$Reverse != "+",]
```

Extract sequence from Phospho (RHSTY) probability column

```{r}
pRHSTY.exp7$Sequence <- gsub("\\(\\d*\\.?\\d*\\)","", pRHSTY.exp7$Phospho..RHSTY..Probabilities)
```

Count number of K or R in sequence and substract 1 in order to get number of missed cleavages

```{r}
pRHSTY.exp7$missed.cleavages <- str_count(pRHSTY.exp7$Sequence, "K|R")-1
```

Split data in three categories: pArg, pHis and pSTY

```{r}
pArg <- pRHSTY.exp7[pRHSTY.exp7$Amino.acid == "R",]  
pArg.classI <- pArg[pArg$Localization.prob >= 0.75,]
pHis <- pRHSTY.exp7[pRHSTY.exp7$Amino.acid == "H",]
pHis.classI <- pHis[pHis$Localization.prob >= 0.75,]
pSTY <- pRHSTY.exp7[pRHSTY.exp7$Amino.acid != "R" & pRHSTY.exp7$Amino.acid != "H",]
pSTY.classI <- pSTY[pSTY$Localization.prob >= 0.75,]
```

Calculate occurrence of 0,1,2 or 3 missed cleavages for pArg, pHis and pSTY

```{r}
#pArg
calculate_missed_cleavages <- function(df, AA){
  
  a <- rbind.data.frame(cbind("missed.cleavges" = 0, "Number"= nrow(df[df$missed.cleavages==0,]), 
                              "Frequency" = round(nrow(df[df$missed.cleavages==0,])/nrow(df)*100,1), 'aa' = AA),
                        cbind("missed.cleavges" = 1, "Number"= nrow(df[df$missed.cleavages==1,]),
                              "Frequency" = round(nrow(df[df$missed.cleavages==1,])/nrow(df)*100,1),'aa' = AA),
                        cbind("missed.cleavges" = 2, "Number"= nrow(df[df$missed.cleavages==2,]),
                              "Frequency" =  round(nrow(df[df$missed.cleavages==2,])/nrow(df)*100,1),'aa' = AA),
                        cbind("missed.cleavges" = 3, "Number"= nrow(df[df$missed.cleavages==3,]),
                              "Frequency" = round(nrow(df[df$missed.cleavages==3,])/nrow(df)*100,1),'aa' = AA))
}

missed.cleavages.pArg <- calculate_missed_cleavages(pArg, "R")
missed.cleavages.pArg.classI <- calculate_missed_cleavages(pArg.classI, "R") 

#pHis
missed.cleavages.pHis <- calculate_missed_cleavages(pHis, "H")
missed.cleavages.pHis.classI <- calculate_missed_cleavages(pHis.classI, "H")
#pSTY

missed.cleavages.pSTY <- calculate_missed_cleavages(pSTY, "STY")
missed.cleavages.pSTY.classI <- calculate_missed_cleavages(pSTY.classI, "STY")

#combine all information in one table
missed.cleavages.overview <- rbind.data.frame(missed.cleavages.pArg, missed.cleavages.pHis, missed.cleavages.pSTY)
missed.cleavages.overview$missed.cleavges <- as.character(missed.cleavages.overview$missed.cleavges)
missed.cleavages.overview$type <- "all"

missed.cleavages.overview.classI <- rbind.data.frame(missed.cleavages.pArg.classI, missed.cleavages.pHis.classI, missed.cleavages.pSTY.classI)
missed.cleavages.overview.classI$missed.cleavges <- as.character(missed.cleavages.overview.classI$missed.cleavges)
missed.cleavages.overview.classI$type <- "classI"


overview <- rbind.data.frame(missed.cleavages.overview, missed.cleavages.overview.classI)
```

Plot distribution of missed cleavages for pHis, pArgg and pSTY

```{r}
pal <- rev(heat_hcl(4,h=c(0,-100), l=c(75,40), c=c(40,80), power=1))

  overview %>%
  mutate(missed.cleavges = fct_relevel(missed.cleavges, "3","2","1", "0" ))%>%
  mutate(Frequency = as.numeric(Frequency)) %>%
  ggplot(aes(x=aa, y=Frequency, fill=missed.cleavges))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = pal)+
  theme_classic()+
  ylab("Number of peptides [%]")+
  xlab("Number of missed cleavages")+
  facet_wrap(~type)


```