---
title: "pArg stability analysis"
author: "Nadine Prust"
date: "4/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Stability analysis of arginine phosphorylation

All necessary data is deposited on Pride and can be downloaded. The used files for this script can be found in the MQ_output_txt.zip file. Add the MQ_output_txt folder then to your project folder to access the required data. 

```{r}
library(ggpubr)
library(ggplot2)
library(ggforce)
library(dplyr)
library(tidyverse)
library(RColorBrewer)
library(reshape2)
```

## Stability at pH2, 4C

Read "evidence.txt" and "msms.txt" from MaxQuant v1.6.17.0 from time series

```{r}
Stability.test.evidence <- read.delim("MQ_output_txt/Synthetic_peptides/Stability_4C/txt/evidence.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
Stability.test.msms <- read.delim("MQ_output_txt/Synthetic_peptides/Stability_4C/txt/msms.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")

```

Filter for contaminants

``` {r}
Stability.test.evidence <- Stability.test.evidence[Stability.test.evidence$Potential.contaminant != "+",]

```

Log2 transform the intensity
```{r}
Stability.test.evidence$log2Intensity <- log2(Stability.test.evidence$Intensity)
```
Change column name so that it is the same in msms and evidence: Best.MS.MS corresponds to id in msms

```{r}
colnames(Stability.test.msms)[colnames(Stability.test.msms)=="id"] <- "Best.MS.MS"

```

Filter for "Multi-MSMS" in evidence table

```{r}
Stability.test.Multi.MSMS <- Stability.test.evidence[Stability.test.evidence$Type == "MULTI-MSMS",]
```

Split date into time points

```{r}
timepoint <- split(Stability.test.Multi.MSMS, f= Stability.test.Multi.MSMS$Experiment)
```

Extract the log2 intensity for the individual time points per replicate and average the intensity per time point

```{r}
intensity.calc <- function(timelist, point1, point2,point3, point4, timepoint, R1.name, R2.name, R3.name, R4.name){
  
  R1 <- as.data.frame(timelist[[point1]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R1 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  
  R1 <- unique(R1)
  
  R2 <- as.data.frame(timelist[[point2]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R2 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  
  R2 <- unique(R2)
  
  R3 <- as.data.frame(timelist[[point3]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R3 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  R3 <- unique(R3)
  
  R4 <- as.data.frame(timelist[[point4]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R4 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  R4 <- unique(R4)
  
  a <- full_join(R1, R2, by="Modified.sequence")
  a<- full_join(a, R3, by="Modified.sequence")
  a <- full_join(a, R4, by="Modified.sequence")
  
  a$Peptide.x <- ifelse(is.na(a$Peptide.x),a$Peptide.y, a$Peptide.x )
  a$Peptide.x <- ifelse(is.na(a$Peptide.x),a$Peptide.x.x, a$Peptide.x )
  a$Peptide.x <- ifelse(is.na(a$Peptide.x),a$Peptide.y.y, a$Peptide.x )
  
  
  
  a$Timepoint.x <- ifelse(is.na(a$Timepoint.x),a$Timepoint.y, a$Timepoint.x )
  a$Timepoint.x <- ifelse(is.na(a$Timepoint.x),a$Timepoint.x.x, a$Timepoint.x )
  a$Timepoint.x <- ifelse(is.na(a$Timepoint.x),a$Timepoint.y.y, a$Timepoint.x )
  
  a$Modifications.x <- ifelse(is.na(a$Modifications.x),a$Modifications.y, a$Modifications.x )
  a$Modifications.x <- ifelse(is.na(a$Modifications.x),a$Modifications.x.x, a$Modifications.x )
  a$Modifications.x <- ifelse(is.na(a$Modifications.x),a$Modifications.y.y, a$Modifications.x )
  
  
  a <- a[, c("Modified.sequence", "Modifications.x", 'R1', 'R2', "R3", "R4", "Timepoint.x", "Peptide.x")]
  
  colnames(a) <- c("Modified.sequence", "Modification", R1.name, R2.name , R3.name, R4.name, "Timepoint", "Peptide")
  return(a)
}

#run function to extract intensities per time point, run it for each time point

timepoint.1 <- intensity.calc(timepoint, "0min_R1", "0min_R2", "0min_R3", "0min_R4", "0min", 
                              "0 min R1", "0 min R2", "0 min R3", "0 min R4")

timepoint.2 <- intensity.calc(timepoint, "15min_R1", "15min_R2", "15min_R3", "15min_R4", "15min", 
                              "15min R1", "15min R2", "15min R3", "15min R4")

timepoint.3 <- intensity.calc(timepoint, "30min_R1", "30min_R2", "30min_R3", "30min_R4", "30min", 
                              "30min R1", "30min R2", "30min R3", "30min R4")

timepoint.4 <- intensity.calc(timepoint, "60min_R1", "60min_R2", "60min_R3", "60min_R4", "60min", 
                              "60min R1", "60min R2", "60min R3", "60min R4")

timepoint.5 <- intensity.calc(timepoint, "120min_R1", "120min_R2", "120min_R3", "120min_R4", "120min", 
                            "120min R1", "120min R2", "120min R3", "120min R4")
#combine all 5 time points
intensity.overview <- full_join(timepoint.1, timepoint.2, by="Modified.sequence")
intensity.overview <- full_join(intensity.overview, timepoint.3, by="Modified.sequence")
intensity.overview <- full_join(intensity.overview, timepoint.4, by="Modified.sequence")
intensity.overview <- full_join(intensity.overview, timepoint.5, by="Modified.sequence")

#add fill up missing infromation for Peptide, Modification and timepoint from all 5 timepoints

intensity.overview$Peptide <- ifelse(is.na(intensity.overview$Peptide),
                                     intensity.overview$Peptide.y.y, 
                                     intensity.overview$Peptide )
intensity.overview$Peptide <- ifelse(is.na(intensity.overview$Peptide),
                                     intensity.overview$Peptide.y, 
                                     intensity.overview$Peptide )

intensity.overview$Peptide <- ifelse(is.na(intensity.overview$Peptide),
                                     intensity.overview$Peptide.x, 
                                     intensity.overview$Peptide )

intensity.overview$Peptide <- ifelse(is.na(intensity.overview$Peptide),
                                     intensity.overview$Peptide.x.x, 
                                     intensity.overview$Peptide )



intensity.overview$Modification <- ifelse(is.na(intensity.overview$Modification),
                                          intensity.overview$Modification.y.y, 
                                          intensity.overview$Modification )

intensity.overview$Modification <- ifelse(is.na(intensity.overview$Modification),
                                          intensity.overview$Modification.y, 
                                          intensity.overview$Modification )

intensity.overview$Modification <- ifelse(is.na(intensity.overview$Modification),
                                          intensity.overview$Modification.x, 
                                          intensity.overview$Modification )

intensity.overview$Modification <- ifelse(is.na(intensity.overview$Modification),
                                          intensity.overview$Modification.x.x, 
                                          intensity.overview$Modification )




intensity.overview$Timepoint <- ifelse(is.na(intensity.overview$Timepoint),
                                          intensity.overview$Timepoint.y.y, 
                                          intensity.overview$Timepoint )
intensity.overview$Timepoint <- ifelse(is.na(intensity.overview$Timepoint),
                                       intensity.overview$Timepoint.y, 
                                       intensity.overview$Timepoint )
intensity.overview$Timepoint <- ifelse(is.na(intensity.overview$Timepoint),
                                       intensity.overview$Timepoint.x, 
                                       intensity.overview$Timepoint )
intensity.overview$Timepoint <- ifelse(is.na(intensity.overview$Timepoint),
                                       intensity.overview$Timepoint.x.x, 
                                       intensity.overview$Timepoint )

#extract necessary columns

intensity.overview.short <- intensity.overview[, c("Modified.sequence", "Peptide", "Modification", 
                                             "0 min R1", "0 min R2", "0 min R3", "0 min R4",
                                             "15min R1", "15min R2", "15min R3", "15min R4",
                                             "30min R1", "30min R2", "30min R3", "30min R4",
                                             "60min R1", "60min R2", "60min R3", "60min R4",
                                             "120min R1", "120min R2", "120min R3", "120min R4")]

# calculate avg and Sd

min0 <- intensity.overview.short[,4:7] %>% mutate(avg.0min = apply(.[1:4], 1, mean, na.rm=TRUE), stDev.0min = apply(.[1:4],1,sd, na.rm=TRUE))
min15 <- intensity.overview.short[,8:11] %>% mutate(avg.15min = apply(.[1:4], 1, mean, na.rm=TRUE), stDev.15min = apply(.[1:4],1,sd, na.rm=TRUE))
min30 <- intensity.overview.short[,12:15] %>% mutate(avg.30min = apply(.[1:4], 1, mean, na.rm=TRUE), stDev.30min = apply(.[1:4],1,sd, na.rm=TRUE))
min60 <- intensity.overview.short[,16:19] %>% mutate(avg.60min = apply(.[1:4], 1, mean, na.rm=TRUE), stDev.60min = apply(.[1:4],1,sd, na.rm=TRUE))
min120 <- intensity.overview.short[,20:23] %>% mutate(avg.120min = apply(.[1:4], 1, mean, na.rm=TRUE), stDev.120min = apply(.[1:4],1,sd, na.rm=TRUE))


#add average and sd to intensity overview
intensity.overview.avg <- cbind.data.frame(intensity.overview.short, min0[,5:6], min15[,5:6], min30[,5:6], min60[,5:6], min120[,5:6])

intensity.overview.avg$AA <- ifelse(intensity.overview.avg$Modification == "Phospho (RHSTY)", 
                                           sub("\\(Phospho .*", "", intensity.overview.avg$Modified.sequence), NA)

intensity.overview.avg$AA <- str_sub(intensity.overview.avg$AA, -1)

#mutant data for the plot
avg.intensity <- intensity.overview.avg[, c("Modified.sequence", "Peptide", "Modification", "AA", "avg.0min", "avg.15min", "avg.30min", "avg.60min", "avg.120min",
                                        "stDev.0min", "stDev.15min", "stDev.30min", "stDev.60min", "stDev.120min")]

avg.intensity.long <- melt(avg.intensity, id= c("Modified.sequence", "Peptide", "Modification", "AA"))
#create SD column
avg.intensity.long$SD <- avg.intensity.long[216:430, 6]
#remove SD rows, that are now in SD column
avg.intensity.long <- avg.intensity.long[-c(216:430),]

avg.intensity.long$variable <- gsub("avg\\.", "", avg.intensity.long$variable)


avg.intensity.long$variable <- factor(avg.intensity.long$variable, 
                                                 levels = c("0min", "15min", "30min", "60min", "120min"))

avg.intensity.long <- avg.intensity.long[!is.na(avg.intensity.long$value),]
#extract information of unmodified peptides and pR peptides (doubly phosphorylated or oxidized peptodes were not taken into account)
avg.intensity.long.pR.unmodi <- avg.intensity.long[avg.intensity.long$Modification == "Phospho (RHSTY)" |
                                                     avg.intensity.long$Modification == "Unmodified",]

avg.intensity.long.pR.unmodi <- avg.intensity.long.pR.unmodi[avg.intensity.long.pR.unmodi$Peptide != "Peptide_04",]

```

Plot pArg stability at pH2, 4C

```{r}
colourCount.Int = length(unique(avg.intensity.long.pR.unmodi$Peptide))
getPalette.Int = colorRampPalette(brewer.pal(9, "Paired"))

Intensity.plot <-
ggplot(avg.intensity.long.pR.unmodi, aes(x=variable, y=value, group= Modified.sequence, color=Peptide))+
  geom_line()+
  geom_point() +
  geom_errorbar(aes(ymin=value-SD, ymax=value+SD), width=.1)+
  scale_color_manual(values = getPalette.Int(colourCount.Int))+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line(), legend.position = "none")+
  facet_wrap(~Modification)+
  ggtitle("pH2, 4°C")
```

Calculate the ratio of pArg/unmodified peptides

```{r}
#extract pArg peptides
avg.pR <- avg.intensity[which(avg.intensity$AA == "R"), ]
#extract unmodified peptides
avg.unmodi <- avg.intensity[which(avg.intensity$Modification == "Unmodified"),]
#combine both in one table next to each other
ratio <- left_join(avg.pR, avg.unmodi, by="Peptide")

#calculate the ratio

ratio.dat <- cbind.data.frame(ratio[,1:4], 
                              ratio$avg.0min.x-ratio$avg.0min.y, 
                              ratio$avg.15min.x-ratio$avg.15min.y,
                              ratio$avg.30min.x-ratio$avg.30min.y,
                              ratio$avg.60min.x-ratio$avg.60min.y,
                              ratio$avg.120min.x-ratio$avg.120min.y)
colnames(ratio.dat) <- c("Modified.sequence", "Peptide", "Modification", "AA", "0min", "15min", "30min", 
                         "60min", "120min")
#mutat date to be plotted
ratio.data.long <- melt(ratio.dat, id= c("Modified.sequence", "Peptide", "Modification", "AA") )
ratio.data.long$variable <- factor(ratio.data.long$variable, 
                                   levels = c("0min", "15min", "30min", "60min", "120min"))

```

Plot ratio at pH2, 4C

```{r}
colourCount.ratio = length(unique(ratio.data.long$Peptide))
getPalette.ratio = colorRampPalette(brewer.pal(9, "Paired"))
Ratio.plot <-
  ggplot(ratio.data.long, aes(x=variable, y=value,group=Peptide, color=Peptide))+
  geom_point(size=2) +
  geom_line(size=0.5)+
  scale_color_manual(values = getPalette.ratio(colourCount.ratio))+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line(), legend.position = "none")+
  ylab("FC(pArg-peptide/unmodified peptide)")+
  ggtitle("pH2, 4°C")
```

## Stability at pH2 room temperature

Read "evidence.txt" and "msms.txt" from overnight stability (were search in MQ (v1.5.30.3) due to ETD,EThcD and HCD fragmentation)

```{r}

Stability.test.evidence.on <- read.delim("MQ_output_txt/Synthetic_peptides/Stability_room_temp/txt/evidence.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
Stability.test.msms.on <- read.delim("MQ_output_txt/Synthetic_peptides/Stability_room_temp/txt/msms.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
```

Filter for contaminants

```{r}
Stability.test.evidence.on <- Stability.test.evidence.on[Stability.test.evidence.on$Potential.contaminant != "+",]
```
Log2 transform intensity

```{r}
Stability.test.evidence.on$log2Intensity <- log2(Stability.test.evidence.on$Intensity)
```

Change column name so that it is the same name in msms and evidence 

```{r}
colnames(Stability.test.msms.on)[colnames(Stability.test.msms.on)=="id"] <- "Best.MS.MS"
```

Select only "Multi-MSMS" in evidence table and "HCD" fragmentation in msms.txt

```{r}
Stability.test.Multi.MSMS.on <- Stability.test.evidence.on[Stability.test.evidence.on$Type == "MULTI-MSMS",]

Stability.test.msms.HCD <- Stability.test.msms.on[Stability.test.msms.on$Fragmentation == "HCD",]
```

Split data into time points

```{r}
timepoint.on <- split(Stability.test.Multi.MSMS.on, f= Stability.test.Multi.MSMS.on$Experiment)
```

Extract the log2 intensity for the individule time points per replicate and average the intensity per time point

```{r}
intensity.calc.on <- function(timelist, point1, point2,point3, point4, timepoint, R1.name, R2.name, R3.name, R4.name){
  
  R1 <- as.data.frame(timelist[[point1]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R1 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  
  R1 <- unique(R1)
  
  R2 <- as.data.frame(timelist[[point2]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R2 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  
  R2 <- unique(R2)
  
  R3 <- as.data.frame(timelist[[point3]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R3 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  R3 <- unique(R3)
  
  R4 <- as.data.frame(timelist[[point4]]) %>% select(Modified.sequence, Modifications, log2Intensity, Proteins) %>% 
    group_by(Modified.sequence, Modifications) %>% summarise(R4 = sum(log2Intensity), Timepoint = timepoint, Peptide = Proteins)
  R4 <- unique(R4)
  
  a <- full_join(R1, R2, by="Modified.sequence")
  a<- full_join(a, R3, by="Modified.sequence")
  a <- full_join(a, R4, by="Modified.sequence")
  
  a$Peptide.x <- ifelse(is.na(a$Peptide.x),a$Peptide.y, a$Peptide.x )
  a$Peptide.x <- ifelse(is.na(a$Peptide.x),a$Peptide.x.x, a$Peptide.x )
  a$Peptide.x <- ifelse(is.na(a$Peptide.x),a$Peptide.y.y, a$Peptide.x )
  
  
  
  a$Timepoint.x <- ifelse(is.na(a$Timepoint.x),a$Timepoint.y, a$Timepoint.x )
  a$Timepoint.x <- ifelse(is.na(a$Timepoint.x),a$Timepoint.x.x, a$Timepoint.x )
  a$Timepoint.x <- ifelse(is.na(a$Timepoint.x),a$Timepoint.y.y, a$Timepoint.x )
  
  a$Modifications.x <- ifelse(is.na(a$Modifications.x),a$Modifications.y, a$Modifications.x )
  a$Modifications.x <- ifelse(is.na(a$Modifications.x),a$Modifications.x.x, a$Modifications.x )
  a$Modifications.x <- ifelse(is.na(a$Modifications.x),a$Modifications.y.y, a$Modifications.x )
  
  
  a <- a[, c("Modified.sequence", "Modifications.x", 'R1', 'R2', "R3", "R4", "Timepoint.x", "Peptide.x")]
  
  colnames(a) <- c("Modified.sequence", "Modification", R1.name, R2.name , R3.name, R4.name, "Timepoint", "Peptide")
  return(a)
}

timepoint.on.1 <- intensity.calc.on(timepoint.on, "0min_R1", "0min_R2", "0min_R3", "0min_R4", "0min", 
                                 "0min R1", "0min R2", "0min R3", "0min R4")
timepoint.on.2 <- intensity.calc.on(timepoint.on, "24hr_R1", "24hr_R2", "24hr_R3", "24hr_R4", "24hr", 
                                 "24hr R1", "24hr R2", "24hr R3", "24hr R4")
#combine both timepoints
intensity.overview.on <- full_join(timepoint.on.1, timepoint.on.2, by="Modified.sequence")

#add missing information of Peptide, modification and timepoint
intensity.overview.on$Peptide.x <- ifelse(is.na(intensity.overview.on$Peptide.x),
                                          intensity.overview.on$Peptide.y, 
                                          intensity.overview.on$Peptide.x )


intensity.overview.on$Modification.x <- ifelse(is.na(intensity.overview.on$Modification.x),
                                               intensity.overview.on$Modification.y, 
                                               intensity.overview.on$Modification.x )

intensity.overview.on$Timepoint.x <- ifelse(is.na(intensity.overview.on$Timepoint.x),
                                            intensity.overview.on$Timepoint.y, 
                                            intensity.overview.on$Timepoint.x )

#extract neccessary columns

intensity.overview.on.short <- intensity.overview.on[, c("Modified.sequence", "Peptide.x", "Modification.x", 
                                                         "0min R1", "0min R2", "0min R3", "0min R4",
                                                         "24hr R1", "24hr R2", "24hr R3", "24hr R4")]

### calculate avg and Sd
hr0 <- intensity.overview.on.short[,4:7] %>% mutate(avg.0min = apply(.[1:4], 1, mean, na.rm=TRUE), stDev.0min = apply(.[1:4],1,sd, na.rm=TRUE))
hr24 <- intensity.overview.on.short[,8:11] %>% mutate(avg.24hr = apply(.[1:4], 1, mean, na.rm=TRUE), stDev.24hr = apply(.[1:4],1,sd, na.rm=TRUE))

#add average and standard deviation to intensity overview
intensity.overview.avg.on <- cbind.data.frame(intensity.overview.on.short, hr0[,5:6], hr24[,5:6])

intensity.overview.avg.on$AA <- ifelse(intensity.overview.avg.on$Modification == "Phospho (RHSTY)", 
                                       sub("\\(ph.*", "", intensity.overview.avg.on$Modified.sequence), NA)

intensity.overview.avg.on$AA <- str_sub(intensity.overview.avg.on$AA, -1)

#extract neccessary columns
avg.intensity.on <- intensity.overview.avg.on[, c("Modified.sequence", "Peptide.x", "Modification.x", "AA", "avg.0min", "avg.24hr",
                                                  "stDev.0min", "stDev.24hr")]

```

Calculate the ratio of pArg/unmodified peptides

```{r}
#extract pArg peptides
avg.pR.on <- avg.intensity.on[which(avg.intensity.on$AA == "R"), ]
#extract unmodified peptides
avg.unmodi.on <- avg.intensity.on[which(avg.intensity.on$Modification == "Unmodified"),]
#combine both list next to each other
ratio.on <- left_join(avg.pR.on, avg.unmodi.on, by="Peptide.x")

#calculate the ratio
ratio.dat.on <- cbind.data.frame(ratio.on[,1:4], 
                                 ratio.on$avg.0min.x-ratio.on$avg.0min.y,
                                 ratio.on$avg.24hr.x-ratio.on$avg.24hr.y)
colnames(ratio.dat.on) <- c("Modified.sequence", "Peptide", "Modification", "AA", "0min", "24hr")

#mutat data to be plotted
ratio.data.on.long <- melt(ratio.dat.on, id= c("Modified.sequence", "Peptide", "Modification", "AA") )
ratio.data.on.long$variable <- factor(ratio.data.on.long$variable, 
                                      levels = c("0min", "24hr"))


```

Plot ratio at pH2, room temperature

```{r}
Ratio.plot.on <-
  ggplot(ratio.data.on.long, aes(x=variable, y=value,group=Peptide, color=Peptide))+
  geom_point(size=2) +
  geom_line(size=0.5)+
  scale_color_manual(values = getPalette.ratio(colourCount.ratio))+
  theme(axis.title.x = element_blank(), panel.background =  element_blank(), axis.line = element_line(), legend.position = "none")+
  ylab("FC(pArg-peptide/unmodified peptide)")+
  ggtitle("pH2, room temperature")
```

## Retention time (RT) analysis

Read "evidence.txt" and "msms.txt" from iRT runs of synthetic pArg and pSTY peptides

```{r}
evidence.pArg <- read.delim("MQ_output_txt/Synthetic_peptides/Retentiontime/pArg/txt/evidence_pArg_iRT.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
msms.pArg <- read.delim("MQ_output_txt/Synthetic_peptides/Retentiontime/pArg/txt/msms_pArg_iRT.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
evidence.pSTY <- read.delim("MQ_output_txt/Synthetic_peptides/Retentiontime/pSTY/txt/evidence_pSTY_iRT.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")
msms.pSTY <- read.delim("MQ_output_txt/Synthetic_peptides/Retentiontime/pSTY/txt/msms_pSTY_iRT.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")

```

Filter for contaminants 

```{r}
evidence.pArg <- evidence.pArg[evidence.pArg$Potential.contaminant != "+",]

evidence.pSTY <- evidence.pSTY[evidence.pSTY$Potential.contaminant != "+",]
```

Log2 tranfomration of intensities

```{r}
evidence.pArg$log2Intensity <- log2(evidence.pArg$Intensity)
evidence.pSTY$log2Intensity <- log2(evidence.pSTY$Intensity)

```

Change column name so that it is the same in msms.txt and evidence.txt

```{r}
colnames(msms.pArg)[colnames(msms.pArg)=="id"] <- "Best.MS.MS"
colnames(msms.pSTY)[colnames(msms.pSTY)=="id"] <- "Best.MS.MS"

```

Filter for "MULTI-MSMS" in evidence table

```{r}
evidence.pArg.Multi.MSMS <- evidence.pArg[evidence.pArg$Type == "MULTI-MSMS",]
evidence.pSTY.Multi.MSMS <- evidence.pSTY[evidence.pSTY$Type == "MULTI-MSMS",]

```

Extract site of phosphorylation

```{r}
evidence.pArg.Multi.MSMS$AA <-  ifelse(evidence.pArg.Multi.MSMS$Modification == "Phospho (RHSTY)", 
                                      sub("\\(Phospho .*", "", evidence.pArg.Multi.MSMS$Modified.sequence), NA)

evidence.pArg.Multi.MSMS$AA <- str_sub(evidence.pArg.Multi.MSMS$AA, -1)

evidence.pSTY.Multi.MSMS$AA <-  ifelse(evidence.pSTY.Multi.MSMS$Modification == "Phospho (RHSTY)", 
                                       sub("\\(Phospho .*", "", evidence.pSTY.Multi.MSMS$Modified.sequence), NA)

evidence.pSTY.Multi.MSMS$AA <- str_sub(evidence.pSTY.Multi.MSMS$AA, -1)

```

Extract best scoring peptides for cases of multiple identification

```{r}
RT.pArg <- evidence.pArg.Multi.MSMS %>% select(Modified.sequence, Modifications, Proteins, Retention.time, AA, Charge, Score, Experiment) %>%
  group_by(Modified.sequence, Modifications, AA, Charge, Experiment) %>% 
  slice_max(Score)

RT.pSTY <- evidence.pSTY.Multi.MSMS %>% select(Modified.sequence, Modifications, Proteins, Retention.time, AA, Charge, Score, Experiment) %>%
  group_by(Modified.sequence, Modifications, AA, Charge, Experiment) %>% 
  slice_max(Score)
```

Calculate the average RT and SD based on Sequence, Modification, AA and Charge

```{r}
RT.pArg.avg <- RT.pArg %>% select(Modified.sequence, Modifications, Proteins, Retention.time, AA, Charge) %>%
  group_by(Modified.sequence, Modifications, AA, Charge) %>% 
  mutate(avg.RT = mean(Retention.time), sd.RT = sd(Retention.time)) %>% distinct(avg.RT,.keep_all = TRUE)

RT.pArg.avg$ID <- paste(RT.pArg.avg$Proteins, RT.pArg.avg$Charge, sep = ";")

RT.pSTY.avg <- RT.pSTY %>% select(Modified.sequence, Modifications, Proteins, Retention.time, AA, Charge) %>%
  group_by(Modified.sequence, Modifications, AA, Charge) %>% 
  mutate(avg.RT = mean(Retention.time), sd.RT = sd(Retention.time)) %>% distinct(avg.RT,.keep_all = TRUE)
RT.pSTY.avg$ID <- paste(RT.pSTY.avg$Proteins, RT.pSTY.avg$Charge, sep = ";")
###Extract RT information for unmodified sequences
RT.pArg.unmodified <- RT.pArg.avg[RT.pArg.avg$Modifications == "Unmodified",]
RT.pSTY.unmodified <- RT.pSTY.avg[RT.pSTY.avg$Modifications == "Unmodified",]
###extract RT information for pArg
RT.pArg.R <- RT.pArg.avg[RT.pArg.avg$Modifications == "Phospho (RHSTY)" & RT.pArg.avg$AA == "R",]

RT.pSTY <- RT.pSTY.avg[RT.pSTY.avg$Modifications == "Phospho (RHSTY)" & RT.pSTY.avg$AA == "S" |
                         RT.pSTY.avg$Modifications == "Phospho (RHSTY)" & RT.pSTY.avg$AA == "T"  |
                         RT.pSTY.avg$Modifications == "Phospho (RHSTY)" & RT.pSTY.avg$AA == "Y" ,]
#combin RT for pSTY and pArg based on ID
RT.comparison <- full_join(RT.pSTY, RT.pArg.R, by = "ID")
#filter for valdi values (peptides that are identified as pArg and pSTY)
RT.comparison.valid <- RT.comparison[!is.na(RT.comparison$Modified.sequence.y) & 
                                       !is.na(RT.comparison$Modified.sequence.x),]
```

Extract information from iRT peptides as reference

```{r}
RT.pR.iRT <- RT.pArg.avg[RT.pArg.avg$Proteins == "iRT",]
RT.pSTY.iRT <- RT.pSTY.avg[RT.pSTY.avg$Proteins == "iRT",]
#all iRT peptides are called iRT change it to counted iRT peptides
RT.pSTY.iRT$Proteins <- c("iRT_01", "iRT_02", "iRT_03", "iRT_04", "iRT_05", "iRT_06",
                          "iRT_07", "iRT_08", "iRT_09", "iRT_10")
RT.pR.iRT$Proteins <- c("iRT_01", "iRT_02", "iRT_03", "iRT_04", "iRT_05", "iRT_06",
                        "iRT_07", "iRT_08", "iRT_09", "iRT_10")


iRT.comparison <- full_join(RT.pSTY.iRT, RT.pR.iRT, by= "Proteins")
```

Plot pSTY as function of pArg RT with iRT as reference standard

```{r}
colourCount.RT = length(unique(RT.comparison.valid$Proteins.x))
getPalette.RT = colorRampPalette(brewer.pal(9, "Paired"))

#plot pARg agains pSTY and regession through iRT peptides colourd in purple
RT.plot.STY <- 
  ggplot(RT.comparison.valid, aes(x=avg.RT.x ,y=avg.RT.y))+
  geom_point(aes(shape=AA.x))+
  geom_errorbar(aes(ymin=avg.RT.y-sd.RT.y, ymax=avg.RT.y+sd.RT.y), width=.025, color="black")+
  geom_errorbar(aes(xmin=avg.RT.x-sd.RT.x, xmax=avg.RT.x+sd.RT.x), width=.025, color="black")+
  geom_point(data = iRT.comparison, aes(x=avg.RT.x ,y=avg.RT.y), color= "purple")+
  geom_smooth(data = iRT.comparison, method = "lm", fullrange=TRUE, color="grey", 
              linetype="dashed")+
  xlim(10,30)+
  ylim(10,30)+
  xlab("Retention time pSTY peptide")+
  ylab("Retention time pArg peptide")+
  theme(panel.background =  element_blank(), axis.line = element_line())

```

Mutate data to plot pARg against unmodified peptides

```{r}

RT.unmdoified.comparison <- full_join(RT.pArg.unmodified, RT.pArg.R, by = "Proteins")
#remove NAs for peptides that are only identified as unmodified or pArg
RT.unmdoified.comparison <- RT.unmdoified.comparison[!is.na(RT.unmdoified.comparison$Experiment.y),]
```

Plot pArg RT as function of unmodified peptide RT 

```{r}
colourCount.RT = length(unique(RT.unmdoified.comparison$Proteins))
getPalette.RT = colorRampPalette(brewer.pal(9, "Paired"))
#plot RT of pArg agains RT unmodified with diagonal for ideal RT behaviour if both would have the same
RT.plot <- 
  ggplot(RT.unmdoified.comparison, aes(x=avg.RT.x ,y=avg.RT.y, color=Proteins))+
  geom_point()+
  geom_errorbar(aes(ymin=avg.RT.y-sd.RT.y, ymax=avg.RT.y+sd.RT.y), width=.025, color="black")+
  geom_errorbar(aes(xmin=avg.RT.x-sd.RT.x, xmax=avg.RT.x+sd.RT.x), width=.025, color="black")+
  scale_color_manual(values = getPalette.Int(colourCount.Int))+
  xlim(10,30)+
  ylim(10,30)+
  geom_abline(intercept = 0, slope = 1, color="black", 
              linetype="dashed", size=1)+
  xlab("Retention time unmodified peptide")+
  ylab("Retention time pArg peptide")+
  theme(panel.background =  element_blank(), axis.line = element_line(), legend.position = "none")
```

## Combine all figures as figure 2

```{r}
ggarrange(ggarrange(Intensity.plot, labels = "a)"),
          ggarrange(Ratio.plot, Ratio.plot.on ,labels = c("b)", "c)"), ncol = 2),
          ggarrange(RT.plot, RT.plot.STY, labels = c("d)", "e)"), ncol = 2),
          nrow = 3)
```