---
title: "pArg_organism_comparison"
author: "Nadine Prust"
date: "6/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comaprison of pArg occurance in S. aureus, E.coli and B. subtilis

All necessary data is deposited on Pride and can be downloaded. The used files for this script can be found in the MQ_output_txt.zip file. Add the MQ_output_txt folder then to your project folder to access the required data.


```{r}
library(dplyr)
library(ggplot2)
```

Read Phospho (RHSTY)Sites.txt from MaxQuant (v1.6.3.4) from E.coli, B. subtilis and S.aureus Sdr mutants

```{r}
pRHSTY.Ecoli <- read.delim("MQ_output_txt/Endogenous_data/Ecoli/txt/Phospho (RHSTY)Sites_Ecoli.txt", header = TRUE,
                           stringsAsFactors = FALSE, dec = ".")
pRHSTY.Bsub <- read.delim("MQ_output_txt/Endogenous_data/Bsubtilis/txt/Phospho (RHSTY)Sites_sub.txt", header = TRUE,
                           stringsAsFactors = FALSE, dec = ".")
phospho.SdrE <- read.delim("MQ_output_txt/Endogenous_data/SdrE/txt/Phospho (RHSTY)Sites_SdrE.txt", header = TRUE, stringsAsFactors = FALSE, dec = ".")

```


```{r}
pRHSTY.Ecoli <- pRHSTY.Ecoli[pRHSTY.Ecoli$Potential.contaminant != "+",]
pRHSTY.Ecoli <- pRHSTY.Ecoli[pRHSTY.Ecoli$Reverse != "+", ]
classI.Ecoli <- pRHSTY.Ecoli[pRHSTY.Ecoli$Localization.prob >= 0.75,]

pRHSTY.Bsub <- pRHSTY.Bsub[pRHSTY.Bsub$Potential.contaminant != "+",]
pRHSTY.Bsub <- pRHSTY.Bsub[pRHSTY.Bsub$Reverse != "+", ]
classI.Bsub <- pRHSTY.Bsub[pRHSTY.Bsub$Localization.prob >= 0.75,]

phospho.SdrE <- phospho.SdrE[phospho.SdrE$Reverse != "+",]
phospho.SdrE <- phospho.SdrE[phospho.SdrE$Potential.contaminant != "+",]

class1.SdrE <- phospho.SdrE[phospho.SdrE$Localization.prob >=0.75,]
```

Calculate distribution of p-sites for the diffrent organisms

```{r}
##Ecoli

#count number of pSTYH w/o localization cutoff
Ecoli.distribution <- data.frame(
  category=c("pS", "pT", "pY", "pH", "pR"),
  count=c(nrow(pRHSTY.Ecoli[pRHSTY.Ecoli$Amino.acid == "S",]), nrow(pRHSTY.Ecoli[pRHSTY.Ecoli$Amino.acid == "T",]), 
          nrow(pRHSTY.Ecoli[pRHSTY.Ecoli$Amino.acid == "Y",]), nrow(pRHSTY.Ecoli[pRHSTY.Ecoli$Amino.acid == "H",]),
          nrow(pRHSTY.Ecoli[pRHSTY.Ecoli$Amino.acid == "R",]))
)
#set leves on category
Ecoli.distribution$category.ordered <- factor(Ecoli.distribution$category, levels = Ecoli.distribution$category)

# calculate percentages

Ecoli.distribution$fraction <- Ecoli.distribution$count / sum(Ecoli.distribution$count)

# Calculate the cumulative percentages (top of each rectangle)
Ecoli.distribution$ymax <- cumsum(Ecoli.distribution$fraction)

# Compute the bottom of each rectangle
Ecoli.distribution$ymin <- c(0, head(Ecoli.distribution$ymax, n=-1))

# Compute label position
Ecoli.distribution$labelPosition <- (Ecoli.distribution$ymax + Ecoli.distribution$ymin) / 2
#compute column with % wih 1 decimal place
Ecoli.distribution$pc <- round(Ecoli.distribution$fraction*100, digits = 1)

#Compute a good label
Ecoli.distribution$label <- paste0(Ecoli.distribution$category, "\n ", Ecoli.distribution$count, "\n ", Ecoli.distribution$pc, "%")

Ecoli.distribution$org <- "Ecoli"
Ecoli.distribution$type <- "all"
####Class I
ClassI.Ecoli.distribution <- data.frame(
  category=c("pS", "pT", "pY", "pH", "pR"),
  count=c(nrow(classI.Ecoli[classI.Ecoli$Amino.acid == "S",]), nrow(classI.Ecoli[classI.Ecoli$Amino.acid == "T",]), 
          nrow(classI.Ecoli[classI.Ecoli$Amino.acid == "Y",]), nrow(classI.Ecoli[classI.Ecoli$Amino.acid == "H",]),
          nrow(classI.Ecoli[classI.Ecoli$Amino.acid == "R",]))
)
#set leves on category
ClassI.Ecoli.distribution$category.ordered <- factor(ClassI.Ecoli.distribution$category, levels = ClassI.Ecoli.distribution$category)

# calculate percentages

ClassI.Ecoli.distribution$fraction <- ClassI.Ecoli.distribution$count / sum(ClassI.Ecoli.distribution$count)

# Calculate the cumulative percentages (top of each rectangle)
ClassI.Ecoli.distribution$ymax <- cumsum(ClassI.Ecoli.distribution$fraction)

# Compute the bottom of each rectangle
ClassI.Ecoli.distribution$ymin <- c(0, head(ClassI.Ecoli.distribution$ymax, n=-1))

# Compute label position
ClassI.Ecoli.distribution$labelPosition <- (ClassI.Ecoli.distribution$ymax + ClassI.Ecoli.distribution$ymin) / 2
#compute column with % wih 1 decimal place
ClassI.Ecoli.distribution$pc <- round(ClassI.Ecoli.distribution$fraction*100, digits = 1)

#Compute a good label
ClassI.Ecoli.distribution$label <- paste0(ClassI.Ecoli.distribution$category, "\n ", ClassI.Ecoli.distribution$count, "\n ", ClassI.Ecoli.distribution$pc, "%")

#organism

ClassI.Ecoli.distribution$org <- "Ecoli"
ClassI.Ecoli.distribution$type <- "classI"

##Bsub
Bsub.distribution <- data.frame(
  category=c("pS", "pT", "pY", "pH", "pR"),
  count=c(nrow(pRHSTY.Bsub[pRHSTY.Bsub$Amino.acid == "S",]), nrow(pRHSTY.Bsub[pRHSTY.Bsub$Amino.acid == "T",]), 
          nrow(pRHSTY.Bsub[pRHSTY.Bsub$Amino.acid == "Y",]), nrow(pRHSTY.Bsub[pRHSTY.Bsub$Amino.acid == "H",]),
          nrow(pRHSTY.Bsub[pRHSTY.Bsub$Amino.acid == "R",]))
)
#set leves on category
Bsub.distribution$category.ordered <- factor(Bsub.distribution$category, levels = Bsub.distribution$category)

# calculate percentages

Bsub.distribution$fraction <- Bsub.distribution$count / sum(Bsub.distribution$count)

# Calculate the cumulative percentages (top of each rectangle)
Bsub.distribution$ymax <- cumsum(Bsub.distribution$fraction)

# Compute the bottom of each rectangle
Bsub.distribution$ymin <- c(0, head(Bsub.distribution$ymax, n=-1))

# Compute label position
Bsub.distribution$labelPosition <- (Bsub.distribution$ymax + Bsub.distribution$ymin) / 2
#compute column with % wih 1 decimal place
Bsub.distribution$pc <- round(Bsub.distribution$fraction*100, digits = 1)

#Compute a good label
Bsub.distribution$label <- paste0(Bsub.distribution$category, "\n ", Bsub.distribution$count, "\n ", Bsub.distribution$pc, "%")

Bsub.distribution$org <- "Bsub"
Bsub.distribution$type <- "all"
###class I

ClassI.Bsub.distribution <- data.frame(
category=c("pS", "pT", "pY", "pH", "pR"),
count=c(nrow(classI.Bsub[classI.Bsub$Amino.acid == "S",]), nrow(classI.Bsub[classI.Bsub$Amino.acid == "T",]), 
        nrow(classI.Bsub[classI.Bsub$Amino.acid == "Y",]), nrow(classI.Bsub[classI.Bsub$Amino.acid == "H",]),
        nrow(classI.Bsub[classI.Bsub$Amino.acid == "R",]))
)
#set leves on category
ClassI.Bsub.distribution$category.ordered <- factor(ClassI.Bsub.distribution$category, levels = ClassI.Bsub.distribution$category)

# calculate percentages

ClassI.Bsub.distribution$fraction <- ClassI.Bsub.distribution$count / sum(ClassI.Bsub.distribution$count)

# Calculate the cumulative percentages (top of each rectangle)
ClassI.Bsub.distribution$ymax <- cumsum(ClassI.Bsub.distribution$fraction)

# Compute the bottom of each rectangle
ClassI.Bsub.distribution$ymin <- c(0, head(ClassI.Bsub.distribution$ymax, n=-1))

# Compute label position
ClassI.Bsub.distribution$labelPosition <- (ClassI.Bsub.distribution$ymax + ClassI.Bsub.distribution$ymin) / 2
#compute column with % wih 1 decimal place
ClassI.Bsub.distribution$pc <- round(ClassI.Bsub.distribution$fraction*100, digits = 1)

#Compute a good label
ClassI.Bsub.distribution$label <- paste0(ClassI.Bsub.distribution$category, "\n ", ClassI.Bsub.distribution$count, "\n ", ClassI.Bsub.distribution$pc, "%")

ClassI.Bsub.distribution$org <- "Bsub"
ClassI.Bsub.distribution$type <- "classI"

##S.aureus SdrE

SdrE.distribution <- data.frame(
  category=c("pS", "pT", "pY", "pH", "pR"),
  count=c(nrow(phospho.SdrE[phospho.SdrE$Amino.acid == "S",]), nrow(phospho.SdrE[phospho.SdrE$Amino.acid == "T",]), 
          nrow(phospho.SdrE[phospho.SdrE$Amino.acid == "Y",]), nrow(phospho.SdrE[phospho.SdrE$Amino.acid == "H",]),
          nrow(phospho.SdrE[phospho.SdrE$Amino.acid == "R",]))
)
#set leves on category
SdrE.distribution$category.ordered <- factor(SdrE.distribution$category, levels = SdrE.distribution$category)

# calculate percentages

SdrE.distribution$fraction <- SdrE.distribution$count / sum(SdrE.distribution$count)


# Calculate the cumulative percentages (top of each rectangle)
SdrE.distribution$ymax <- cumsum(SdrE.distribution$fraction)

# Compute the bottom of each rectangle
SdrE.distribution$ymin <- c(0, head(SdrE.distribution$ymax, n=-1))
# Compute label position
SdrE.distribution$labelPosition <- (SdrE.distribution$ymax + SdrE.distribution$ymin) / 2

#compute column with % wih 1 decimal place
SdrE.distribution$pc <- round(SdrE.distribution$fraction*100, digits = 1)

#Compute a good label
SdrE.distribution$label <- paste0(SdrE.distribution$category, "\n ", SdrE.distribution$count, "\n ", SdrE.distribution$pc, "%")
SdrE.distribution$org <- "Saureus"
SdrE.distribution$type <- "all"


###classI
SdrE.class1.distribution <- data.frame(
  category=c("pS", "pT", "pY", "pH", "pR"),
  count=c(nrow(class1.SdrE[class1.SdrE$Amino.acid == "S",]), nrow(class1.SdrE[class1.SdrE$Amino.acid == "T",]), 
          nrow(class1.SdrE[class1.SdrE$Amino.acid == "Y",]), nrow(class1.SdrE[class1.SdrE$Amino.acid == "H",]),
          nrow(class1.SdrE[class1.SdrE$Amino.acid == "R",]))
)
#set leves on category
SdrE.class1.distribution$category.ordered <- factor(SdrE.class1.distribution$category, levels = SdrE.class1.distribution$category)



# calculate percentages

SdrE.class1.distribution$fraction <- SdrE.class1.distribution$count / sum(SdrE.class1.distribution$count)


# Calculate the cumulative percentages (top of each rectangle)
SdrE.class1.distribution$ymax <- cumsum(SdrE.class1.distribution$fraction)

# Compute the bottom of each rectangle
SdrE.class1.distribution$ymin <- c(0, head(SdrE.class1.distribution$ymax, n=-1))
# Compute label position
SdrE.class1.distribution$labelPosition <- (SdrE.class1.distribution$ymax + SdrE.class1.distribution$ymin) / 2

#compute column with % wih 1 decimal place
SdrE.class1.distribution$pc <- round(SdrE.class1.distribution$fraction*100, digits = 1)

#Compute a good label
SdrE.class1.distribution$label <- paste0(SdrE.class1.distribution$category, "\n ", SdrE.class1.distribution$count, "\n ", SdrE.class1.distribution$pc, "%")
SdrE.class1.distribution$org <- "Saureus"
SdrE.class1.distribution$type <- "classI"
```

Plot percentage of pArg sites for the three different organisms 

```{r}

comparison <- rbind.data.frame(SdrE.distribution, Ecoli.distribution, Bsub.distribution,
                               SdrE.class1.distribution, ClassI.Ecoli.distribution, ClassI.Bsub.distribution)
pArg.comparison <- comparison[comparison$category == "pR",]

pArg.comparison %>% 
  ggplot(aes(x=org, y=pc, group=org, fill=org))+
  geom_bar(stat = "identity", position = position_dodge2())+
  scale_fill_brewer(palette = "PuRd")+
  theme_classic()+
  facet_wrap(~type)+
  ylab("pArg sites [%]")
```